### 2016-07-22 code/combine_years.Rmd
Added compressed data files

---
title: "finalproject"
author: "William Paul"
date: "July 14, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(readr)
library(ggplot2)
library(dplyr)
files96 = sapply(paste("hads", c(seq(1985, 2001, by = 2)), ".txt", sep =""), read_csv)
files96 = sapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".txt", sep ="")
files03 = sapply(files, read_csv)

```

### 2016-07-22 code/combine_years.Rmd
Added correct compression, and initial code to combine datasets on modern ones

@@
 library(readr)
-library(ggplot2)
 library(dplyr)
-files96 = sapply(paste("hads", c(seq(1985, 2001, by = 2)), ".txt", sep =""), read_csv)
-files96 = sapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
-files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".txt", sep ="")
-files03 = sapply(files, read_csv)
-
+mutateyears = function(data, year) {
+  mutate(data, dataset_year=year)
+}
+files96 = lapply(paste("hads", c(seq(1985, 2001, by = 2)), ".bz2", sep =""), read_csv)
+files96 = lapply(files96, data.frame, stringsAsFactors= F)
+files96 = lapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
+files96 = mapply(mutateyears, files03, as.list(seq(1985,2001,by=2)))
+files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".bz2", sep ="")
+files03 = lapply(lapply(files03, read_csv), data.frame, stringsAsFactors= F)
+files03 = mapply(mutateyears, files03, as.list(seq(2003,2013,by=2)))
+new2011 = files03[[4]] %>% select(-VCHRMOV)
+c = rbind(do.call(rbind, files03[c(1,2,3,5,6)]), new2011)
 ```

### 2016-07-23 code/combine_years.Rmd
Now works for all years, and can be run from code dir

--- 2016-07-22 code/combine_years.Rmd
@@
 library(readr)
-library(ggplot2)
 library(dplyr)
-files96 = sapply(paste("hads", c(seq(1985, 2001, by = 2)), ".txt", sep =""), read_csv)
-files96 = sapply(files96, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
-files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".txt", sep ="")
-files03 = sapply(files, read_csv)
-
+library(stringr)
+mutateyears = function(data, year) {
+  mutate(data, DATASET_YEAR=year)
+}
+files03 = paste("hads", c(seq(2003, 2013, by = 2)), ".bz2", sep ="")
+files03 = lapply(lapply(files03, read_csv), data.frame, stringsAsFactors= F)
+files03 = mapply(mutateyears, files03, as.list(seq(2003,2013,by=2)))
+new2011 = files03[[4]] %>% select(-VCHRMOV)
+modernfiles = rbind(do.call(rbind, files03[c(1,2,3,5,6)]), new2011)
+files85 = lapply(paste("hads", c(seq(1985, 1995, by = 2)), ".bz2", sep =""), read_csv)
+files85 = lapply(files85, data.frame, stringsAsFactors= F)
+new1995 = files85[[6]] %>% mutate(L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50) %>%  rename("AGE1"=`AGE`, "FMTMETRO3" =`FMTMETRO`, "STATUS"=`ISTATUS`, "METRO3"=`METRO`)
+files85 = lapply(files85[1:5], function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50) %>%  rename("AGE1"=`AGE`, "FMTMetro3" = `FMTMetro`, "STATUS"=`ISTATUS`, "METRO3"=`METRO`)})
+files85 = append(files85, list(new1995))
+files85 = mapply(mutateyears, files85, as.list(seq(1985,1995,by=2)), SIMPLIFY = FALSE)
+files85 = lapply(files85, function(x){setNames(x, (toupper(names(x)) %>% str_replace("^COSTMED(.+)", "COSTMed\\1")))})
+files85 = lapply(files85, select_, .dots=names(modernfiles))
+files95 = lapply(paste("hads", c(seq(1997, 2001, by = 2)), ".bz2", sep =""), read_csv)
+files95 = lapply(files95, data.frame, stringsAsFactors= F)
+files95 = lapply(files95, function(x){mutate(x, L30 = 0.6 * L50, GL30 = 0.6 * GL50, ABL30 = 0.6 * ABL50)})
+files95 = mapply(mutateyears, files95, as.list(seq(1997,2001,by=2)), SIMPLIFY = FALSE)
+files95 = lapply(files95, function(x){setNames(x, (toupper(names(x)) %>% str_replace("^COSTMED(.+)", "COSTMed\\1")))})
+files95 = lapply(files95, select_, .dots=names(modernfiles))
+oldfiles1 = do.call(rbind, files96)
+oldfiles2 = do.call(rbind, files85)
+totalfiles = rbind(rbind(oldfiles1, oldfiles2), modernfiles)
+write_csv(totalfiles, "../data/cleaned_years.csv")
 ```

### 2016-07-24 code/plots.R
Add first plot

@@ -1,0 +1,17 @@
+library(dplyr)
+library(ggplot2)
+
+set.seed(1337)
+# Take a sample from the data to have a lighter and faster dataset
+cleaned_sample <- sample_n(cleaned_years, 10000)
+
+cleaned_sample %>% 
+  group_by(dataset_year) %>% 
+  # Take the median instead of mean to avoid extreme observations skewing the
+  # result.
+  summarise(burden_year = median(burden)) %>% 
+  ggplot(aes(x = dataset_year, y = burden_year)) + 
+  geom_line()
+
+# The housing price bubble can be seen
+## But we want to analyze the effect of falling incomes as well. To be expanded.

### 2016-07-24 code/plots.R
Add shiny server and ui files

@@ -1,17 +1,29 @@
 library(dplyr)
 library(ggplot2)
+library(readr)
 
-set.seed(1337)
 # Take a sample from the data to have a lighter and faster dataset
-cleaned_sample <- sample_n(cleaned_years, 10000)
 
-cleaned_sample %>% 
-  group_by(dataset_year) %>% 
+#set.seed(1337)
+dirty_sample <- read_csv("../data/cleaned_years.csv") %>%
+  sample_n(10000) 
+
+dirty_sample <- read_csv("../data/dirty_sample.csv")
+
+clean_sample <- dirty_sample %>% 
+  setNames(str_to_lower(names(.))) %>%  
+  mutate_if(is.character, NoQuotations) %>%  
+  na_if(-9) %>%
+  mutate(ownrent = factor(ownrent,   
+                          levels = c("1", "2"), 
+                          labels = c("own", "rent")))
+
+clean_sample %>% 
+  group_by(dataset_year, ownrent) %>% 
   # Take the median instead of mean to avoid extreme observations skewing the
   # result.
   summarise(burden_year = median(burden)) %>% 
-  ggplot(aes(x = dataset_year, y = burden_year)) + 
-  geom_line()
+  ggplot(aes(x = dataset_year, y = burden_year, colour = ownrent)) + geom_line()
 
 # The housing price bubble can be seen
 ## But we want to analyze the effect of falling incomes as well. To be expanded.

### 2016-07-24 code/server.R

@@ -1,0 +1,34 @@
+library(ggplot2)
+library(dplyr)
+library(readr)
+library(maps)
+library(ggmap)
+
+dirty_sample <- read_csv("../data/dirty_sample.csv")
+
+clean_sample <- dirty_sample %>% 
+  setNames(str_to_lower(names(.))) %>%  
+  mutate_if(is.character, NoQuotations) %>%  
+  na_if(-9) %>%
+  mutate(ownrent = factor(ownrent,   
+                          levels = c("1", "2"), 
+                          labels = c("own", "rent")))
+
+shinyServer(function(input, output) {
+  
+  output$plot <- renderPlot({ 
+    
+     clean_sample %>%  
+      group_by(dataset_year, ownrent) %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result. 
+      summarise(burden_year = median(burden, na.rm = T)) %>%  
+      ggplot(aes(x = dataset_year,  
+                 y = burden_year,  
+                 colour = ownrent)) +  
+      geom_line()
+
+  })
+  
+})
+

### 2016-07-24 code/ui.R

@@ -1,0 +1,18 @@
+shinyUI(fluidPage(
+
+  titlePanel("Dialects"),
+  
+  sidebarLayout(
+    sidebarPanel(
+      
+      selectInput("sel_var",
+                  label = "Variable",
+                  choices = names(clean_sample))
+      
+      ),
+    
+    mainPanel(
+      plotOutput("plot")
+    )
+  )
+))

### 2016-07-24 code/plots.R
Make shiny app interactive

@@ -1,29 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(readr)
-
-# Take a sample from the data to have a lighter and faster dataset
-
-#set.seed(1337)
-dirty_sample <- read_csv("../data/cleaned_years.csv") %>%
-  sample_n(10000) 
-
-dirty_sample <- read_csv("../data/dirty_sample.csv")
-
-clean_sample <- dirty_sample %>% 
-  setNames(str_to_lower(names(.))) %>%  
-  mutate_if(is.character, NoQuotations) %>%  
-  na_if(-9) %>%
-  mutate(ownrent = factor(ownrent,   
-                          levels = c("1", "2"), 
-                          labels = c("own", "rent")))
-
-clean_sample %>% 
-  group_by(dataset_year, ownrent) %>% 
-  # Take the median instead of mean to avoid extreme observations skewing the
-  # result.
-  summarise(burden_year = median(burden)) %>% 
-  ggplot(aes(x = dataset_year, y = burden_year, colour = ownrent)) + geom_line()
-
-# The housing price bubble can be seen
-## But we want to analyze the effect of falling incomes as well. To be expanded.

### 2016-07-24 code/server.R

@@ -19,13 +19,13 @@ shinyServer(function(input, output) {
   output$plot <- renderPlot({ 
     
      clean_sample %>%  
-      group_by(dataset_year, ownrent) %>%  
+      group_by_("dataset_year", input$sel_var) %>%  
       # Take the median instead of mean to avoid extreme observations skewing the 
       # result. 
       summarise(burden_year = median(burden, na.rm = T)) %>%  
-      ggplot(aes(x = dataset_year,  
-                 y = burden_year,  
-                 colour = ownrent)) +  
+      ggplot(aes_string(x = "dataset_year",   
+                        y = "burden_year",   
+                        colour = input$sel_var)) +  
       geom_line()
 
   })

### 2016-07-24 code/ui.R

@@ -7,7 +7,8 @@ shinyUI(fluidPage(
       
       selectInput("sel_var",
                   label = "Variable",
-                  choices = names(clean_sample))
+                  choices = names(clean_sample),
+                  selected = "ownrent")
       
       ),
     

### 2016-07-24 code/server.R
Add comments to ui and server code

@@ -4,11 +4,16 @@ library(readr)
 library(maps)
 library(ggmap)
 
+# dirty_sample is just a sample of the full combined but unmodified dataset,
+# saved to a separate file for convenience:
+# sample_n(dirty_dataset, 10000)
 dirty_sample <- read_csv("../data/dirty_sample.csv")
 
+# Cleaning the data
 clean_sample <- dirty_sample %>% 
   setNames(str_to_lower(names(.))) %>%  
   mutate_if(is.character, NoQuotations) %>%  
+  # -9 has been used as a marker for NA, replace them with NA
   na_if(-9) %>%
   mutate(ownrent = factor(ownrent,   
                           levels = c("1", "2"), 
@@ -19,10 +24,12 @@ shinyServer(function(input, output) {
   output$plot <- renderPlot({ 
     
      clean_sample %>%  
+      # Use NSE
       group_by_("dataset_year", input$sel_var) %>%  
       # Take the median instead of mean to avoid extreme observations skewing the 
-      # result. 
+      # result.
       summarise(burden_year = median(burden, na.rm = T)) %>%  
+      # Use NSE
       ggplot(aes_string(x = "dataset_year",   
                         y = "burden_year",   
                         colour = input$sel_var)) +  

### 2016-07-24 code/ui.R

@@ -1,13 +1,15 @@
 shinyUI(fluidPage(
 
-  titlePanel("Dialects"),
+  titlePanel("Burden by year and category"),
   
   sidebarLayout(
     sidebarPanel(
       
       selectInput("sel_var",
-                  label = "Variable",
+                  label = "Grouping variable",
                   choices = names(clean_sample),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
                   selected = "ownrent")
       
       ),

### 2016-07-25 code/combine_years.R
Renamed output file, changed format to .R, cleaned up code

@@ -1,0 +1,86 @@
+library(readr)
+library(dplyr)
+library(stringr)
+MutateYears = function(data, year) {
+  # Simply marks the dataset with the year to categorize the original data.
+  mutate(data, 
+         DATASET_YEAR = year)
+  }
+UpperCaseColNames = function(frame) {
+  # Years vary over the years, so this converts the column 
+  # names to the same style as that of the latest dataset.
+  setNames(frame, 
+           (toupper(names(frame)) %>% 
+              str_replace("^COSTMED(.+)",
+                          "COSTMed\\1")))
+  }
+AddMissingCols = function(frame) {
+  # Datasets before 2003 are missing the three columns below, 
+  # so this function adds them to unify all of the datasets
+  mutate(frame, 
+         L30 = 0.6 * L50, 
+         GL30 = 0.6 * GL50, 
+         ABL30 = 0.6 * ABL50)
+  }
+RenameCols = function(frame) {
+  # Naming convention changed after the 1995 dataset, 
+  # so we match the column names.
+  rename(frame, 
+         "AGE1"=`AGE`, 
+         "FMTMETRO3" = `FMTMETRO`, 
+         "STATUS"=`ISTATUS`, 
+         "METRO3"=`METRO`)
+  }
+base_files = paste("../data/hads",          # Loading the most recent datasets 
+                c(seq(2003, 2013, by = 2)), # first to give reference for other
+                ".bz2",                     # two sets.
+                sep ="") %>%  
+  lapply(read_csv) %>% 
+  lapply(data.frame,  # Have to convert to data frames as lapply seems 
+         stringsAsFactors= F) #to recognize csv output as named lists.
+base_files = mapply(MutateYears, # Need to distinguish different years
+                 base_files,     #for datasets, so add column to do just that.
+                 as.list(seq(2003,2013,by=2)))
+modified_year = base_files[[4]] %>% # The 2009 dataset has an extra unspecified
+  select(-VCHRMOV)                  # column that we need to remove.
+modern_files = rbind(do.call(rbind, # Combine all of the recent years into one.
+                            base_files[c(1,2,3,5,6)]), 
+                    modified_year)
+first_decade = paste("../data/hads",             # Load first decade separately 
+                     c(seq(1985, 1995, by = 2)), # to account for missing 
+                     ".bz2",                     # columns and different 
+                     sep ="") %>%                # names.
+  lapply(read_csv) %>% 
[…]

### 2016-07-25 code/server.R
Fix dataset name

@@ -6,7 +6,7 @@ library(ggmap)
 
 # dirty_sample is just a sample of the full combined but unmodified dataset,
 # saved to a separate file for convenience:
-# sample_n(dirty_dataset, 10000)
+# sample_n(combined_years, 10000)
 dirty_sample <- read_csv("../data/dirty_sample.csv")
 
 # Cleaning the data

### 2016-07-25 code/parse_data.R
code

@@ -7,8 +7,17 @@ NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")
 }
 
-y2013 <- read_csv("data/thads2013n.txt") %>%
+clean <- modernfiles %>%
+  na_if(., -9) %>%
   setNames(str_to_lower(names(.))) %>%  
   mutate_if(is.character, NoQuotations)
+
+#plot
+clean %>% 
+  group_by(dataset_year) %>% 
+  summarise(burden_year = mean(burden)) %>% 
+  ggplot(aes(x = dataset_year, y = burden_year)) + 
+  geom_line()
   
+ 
   

### 2016-07-25 code/parse_data.R
clean data

@@ -1,23 +1,21 @@
-library(readr)
-library(dplyr)
-library(stringr)
-library(tidyr)
+# dirty_sample is just a sample of the full combined but unmodified dataset,
+# saved to a separate file for convenience:
+combined_years <- read_csv("../data/combined_years.csv")
+
 
 NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")
 }
 
-clean <- modernfiles %>%
-  na_if(., -9) %>%
+# Cleaning the data
+clean_years <- combined_years %>% 
   setNames(str_to_lower(names(.))) %>%  
-  mutate_if(is.character, NoQuotations)
+  mutate_if(is.character, NoQuotations) %>%  
+  # -9 has been used as a marker for NA, replace them with NA
+  na_if(-9) %>%
+  mutate(ownrent = factor(ownrent,   
+                          levels = c("1", "2"), 
+                          labels = c("own", "rent")))
 
-#plot
-clean %>% 
-  group_by(dataset_year) %>% 
-  summarise(burden_year = mean(burden)) %>% 
-  ggplot(aes(x = dataset_year, y = burden_year)) + 
-  geom_line()
-  
- 
-  
+set.seed(1337)  
+clean_sample <- sample_n(clean_years, 100000)

### 2016-07-25 code/server.R
Remove data cleaning code from server.R

@@ -1,26 +1,13 @@
 library(ggplot2)
 library(dplyr)
 library(readr)
-library(maps)
-library(ggmap)
-
-# dirty_sample is just a sample of the full combined but unmodified dataset,
-# saved to a separate file for convenience:
-# sample_n(combined_years, 10000)
-dirty_sample <- read_csv("../data/dirty_sample.csv")
-
-# Cleaning the data
-clean_sample <- dirty_sample %>% 
-  setNames(str_to_lower(names(.))) %>%  
-  mutate_if(is.character, NoQuotations) %>%  
-  # -9 has been used as a marker for NA, replace them with NA
-  na_if(-9) %>%
-  mutate(ownrent = factor(ownrent,   
-                          levels = c("1", "2"), 
-                          labels = c("own", "rent")))
+library(stringr)
+ 
 
 shinyServer(function(input, output) {
-  
+ 
+clean_sample <- read_csv("../data/clean_sample.csv")
+
   output$plot <- renderPlot({ 
     
      clean_sample %>%  

### 2016-07-25 code/parsedata.R
add new

@@ -1,3 +1,8 @@
+library(readr)
+library(dplyr)
+library(stringr)
+library(tidyr)
+
 # dirty_sample is just a sample of the full combined but unmodified dataset,
 # saved to a separate file for convenience:
 combined_years <- read_csv("../data/combined_years.csv")
@@ -17,5 +22,15 @@ clean_years <- combined_years %>%
                           levels = c("1", "2"), 
                           labels = c("own", "rent")))
 
+#plot
+clean %>% 
+  group_by(dataset_year) %>% 
+  summarise(burden_year = mean(burden)) %>% 
+  ggplot(aes(x = dataset_year, y = burden_year)) + 
+  geom_line()
+  
+ 
+  
 set.seed(1337)  
 clean_sample <- sample_n(clean_years, 100000)
+

### 2016-07-25 code/parse_data.R
new file

@@ -3,11 +3,34 @@ library(dplyr)
 library(stringr)
 library(tidyr)
 
+# dirty_sample is just a sample of the full combined but unmodified dataset,
+# saved to a separate file for convenience:
+combined_years <- read_csv("../data/combined_years.csv")
+
+
 NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")
 }
 
-y2013 <- read_csv("data/thads2013n.txt") %>%
+# Cleaning the data
+clean_years <- combined_years %>% 
   setNames(str_to_lower(names(.))) %>%  
-  mutate_if(is.character, NoQuotations)
+  mutate_if(is.character, NoQuotations) %>%  
+  # -9 has been used as a marker for NA, replace them with NA
+  na_if(-9) %>%
+  mutate(ownrent = factor(ownrent,   
+                          levels = c("1", "2"), 
+                          labels = c("own", "rent")))
+
+#plot
+clean %>% 
+  group_by(dataset_year) %>% 
+  summarise(burden_year = mean(burden)) %>% 
+  ggplot(aes(x = dataset_year, y = burden_year)) + 
+  geom_line()
+  
+ 
+  
+set.seed(1337)  
+clean_sample <- sample_n(clean_years, 100000)
 

### 2016-07-25 code/parse_data.R
removeed plot code

@@ -22,15 +22,6 @@ clean_years <- combined_years %>%
                           levels = c("1", "2"), 
                           labels = c("own", "rent")))
 
-#plot
-clean %>% 
-  group_by(dataset_year) %>% 
-  summarise(burden_year = mean(burden)) %>% 
-  ggplot(aes(x = dataset_year, y = burden_year)) + 
-  geom_line()
-  
- 
-  
 set.seed(1337)  
 clean_sample <- sample_n(clean_years, 100000)
 

### 2016-07-25 code/parse_data.R
write CSVs

@@ -25,3 +25,6 @@ clean_years <- combined_years %>%
 set.seed(1337)  
 clean_sample <- sample_n(clean_years, 100000)
 
+#write csv files
+write_csv(clean_years, path = "../data/clean_years.csv")
+write_csv(clean_sample, path = "../data/clean_sample.csv")

### 2016-07-28 code/parse_data.R
Add manual column types to avoid parsing failures

@@ -3,10 +3,9 @@ library(dplyr)
 library(stringr)
 library(tidyr)
 
-# dirty_sample is just a sample of the full combined but unmodified dataset,
-# saved to a separate file for convenience:
-combined_years <- read_csv("../data/combined_years.csv")
-
+combined_years <- read_csv("../data/combined_years.csv",
+                           # Add col_types manually to avoid parsing failures
+                           col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd")
 
 NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")
@@ -17,14 +16,20 @@ clean_years <- combined_years %>%
   setNames(str_to_lower(names(.))) %>%  
   mutate_if(is.character, NoQuotations) %>%  
   # -9 has been used as a marker for NA, replace them with NA
-  na_if(-9) %>%
+  na_if(-9)  %>%
   mutate(ownrent = factor(ownrent,   
                           levels = c("1", "2"), 
                           labels = c("own", "rent")))
 
-set.seed(1337)  
-clean_sample <- sample_n(clean_years, 100000)
+  clean_sample <- mutate(clean_sample, ownrent = factor(ownrent,   
+                          levels = c("1", "2"), 
+                          labels = c("own", "rent")))
+# Free some memory
+remove(combined_years)
 
-#write csv files
+set.seed(1337)  
+clean_sample <- sample_n(clean_years, 10000)
+
+# Write csv files
 write_csv(clean_years, path = "../data/clean_years.csv")
 write_csv(clean_sample, path = "../data/clean_sample.csv")

### 2016-07-28 code/parse_data.R
Fix issues with variable "ownrent"

@@ -17,18 +17,17 @@ clean_years <- combined_years %>%
   mutate_if(is.character, NoQuotations) %>%  
   # -9 has been used as a marker for NA, replace them with NA
   na_if(-9)  %>%
+  # Clean ownrent
   mutate(ownrent = factor(ownrent,   
                           levels = c("1", "2"), 
                           labels = c("own", "rent")))
 
-  clean_sample <- mutate(clean_sample, ownrent = factor(ownrent,   
-                          levels = c("1", "2"), 
-                          labels = c("own", "rent")))
 # Free some memory
 remove(combined_years)
 
+# Make a sample for faster developing etc.
 set.seed(1337)  
-clean_sample <- sample_n(clean_years, 10000)
+clean_sample <- sample_n(clean_years, 1000)
 
 # Write csv files
 write_csv(clean_years, path = "../data/clean_years.csv")

### 2016-07-28 code/combine_years.R
Added y axis selecting functionality and renamed dataset year to year

@@ -4,7 +4,7 @@ library(stringr)
 MutateYears = function(data, year) {
   # Simply marks the dataset with the year to categorize the original data.
   mutate(data, 
-         DATASET_YEAR = year)
+         YEAR = year)
   }
 UpperCaseColNames = function(frame) {
   # Years vary over the years, so this converts the column 

### 2016-07-28 code/ui.R

@@ -5,12 +5,18 @@ shinyUI(fluidPage(
   sidebarLayout(
     sidebarPanel(
       
-      selectInput("sel_var",
-                  label = "Grouping variable",
+      selectInput("sel_x",
+                  label = "Colour",
                   choices = names(clean_sample),
                   # Select "ownrent" as initial value to avoid getting a 
                   # non-categorical variable by default
-                  selected = "ownrent")
+                  selected = "ownrent"),
+      selectInput("sel_y",
+                  label = "Y axis",
+                  choices = names(clean_sample),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
+                  selected = "burden")
       
       ),
     

### 2016-07-29 code/ui.R
Fix the issue requiring manual loading of data to shiny

@@ -1,3 +1,7 @@
+library(readr)
+
+clean_sample <- read_csv("../data/clean_sample.csv")
+
 shinyUI(fluidPage(
 
   titlePanel("Burden by year and category"),
@@ -17,7 +21,6 @@ shinyUI(fluidPage(
                   # Select "ownrent" as initial value to avoid getting a 
                   # non-categorical variable by default
                   selected = "burden")
-      
       ),
     
     mainPanel(

### 2016-07-30 code/nikolais_plot.Rmd
Add plot of burden and household sizes

```{r}
library(ggplot2)
library(dplyr)
library(readr)
library(stringr)

clean_years <- read_csv("../data/clean_years.csv")

  clean_years %>%
    mutate(per = cut(per, 
                     breaks = c(0, 1, 2, Inf),
                     labels = c("1", "2", "3 or more")))  %>%
    # Use NSE
    group_by(year, per) %>%  
    # Take the median instead of mean to avoid extreme observations skewing the 
    # result.
    summarise(new_y = median(burden, na.rm = T)) %>%
    ggplot(aes(x = year,   
                      y = new_y,   
                      colour = per)) +  
    geom_line() +
    labs(colour = "Household size",
         x = "Year",
         y = "Median burden")
```

The graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013. The median burden varies dramatically between households of different sizes. The highest burden thorought the years has been on households of one person. This is understandable, since people living alone cannot split any costs between other earners. Surprisingly, households of two people have the lowest burdens overall. Our hypothesis is that this is due to a large proportion of two-earner couples in this group. As housing cost burden is simply a household’s monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden. In households of 3 or more people, the burden picks up again. This is likely due to a larger amount of children and / or retirees in larger households.
Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007. The most probable exlanation is the housing market price bubble. Once the crisis hit in 2007, the effect of dropping incomes continued to increase burdens until 2011. Only in 2013 data did the burdens start to fall from the unprecedently high levels.

### 2016-07-30 code/graph_regdat_relincome.R
APL median by region code

@@ -1,0 +1,16 @@
+library(ggplot2)
+library(readr)
+library(stringr)
+library(dplyr)
+col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + geom_line() 

### 2016-07-30 code/nikolais_plot.Rmd
Specify coltypes
Fix typo
Set echo=F

@@
-```{r}
+```{r, echo=F}
 library(ggplot2)
@@
 
-clean_years <- read_csv("../data/clean_years.csv")
+col_types <- "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
 
@@
                      labels = c("1", "2", "3 or more")))  %>%
-    # Use NSE
     group_by(year, per) %>%  
@@
 The median burden varies dramatically between households of different sizes.
-The highest burden thorought the years has been on households of one person.
+The highest burden through the years has been on households of one person.
 This is understandable, since people living alone cannot split any costs between other earners.

### 2016-07-30 code/aplmed_region.Rmd
Add plot Rmd

---
title: "APL_region"
author: "Sriharsha Kondapally"
date: "July 30, 2016"
output: html_document
---

```{r}
library(ggplot2)
library(readr)
library(stringr)
library(dplyr)

col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"

clean_years <- read_csv("../data/clean_years.csv", col_types = col)

clean_years %>%
  group_by(dataset_year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  ggplot(aes(x = dataset_year,   
                  y = new_y,   
                  colour = region)) + geom_line() +
  labs(x = "year", y = "Median Adjusted Income (per persons)", 
       title = "APLmed by region")
```

This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West). IT is clear that the Northeast and the West are the most expensive regions. The median adjusted income increases steadily (roughly linearly) over time. There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.

### 2016-07-30 code/graph_regdat_relincome.R

@@ -13,4 +13,6 @@ clean_years %>%
   summarise(new_y = median(aplmed, na.rm = T)) %>%
   ggplot(aes(x = year,   
                   y = new_y,   
-                  colour = region)) + geom_line() 
+                  colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "APLmed by region")

### 2016-08-01 code/nikolais_plot.Rmd

@@
-```{r, echo=F}
+---
+output: html_document
+---
+```{r, include=F}
 library(ggplot2)
@@
 library(stringr)
+library(xml2)
+library(rvest)
+library(tidyr)
+library(purrr)
+library(maps)
+```
 
+```{r, echo=F}
 col_types <- "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
@@
 This is understandable, since people living alone cannot split any costs between other earners.
+
 Surprisingly, households of two people have the lowest burdens overall.
@@
 This is likely due to a larger amount of children and / or retirees in larger households.
+
 Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007.
@@
 Only in 2013 data did the burdens start to fall from the unprecedently high levels.
+
+```{r, include=F}
+url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
+html <- read_html(url)
+xpath <- '//table[@class="tableizer-table"]'
+
+regions_tbl <- html %>% 
+  html_nodes(xpath = xpath) %>%
+  html_table %>%
+  .[[1]]
+
+region_1 <- regions_tbl[2:7,] %>% 
+  gather(key = region, value = state, 1:2) %>% 
+  slice(1:9) %>%
+  mutate(region = 1)
+
+region_2 <- regions_tbl[10:15,] %>% 
+  gather(key = region, value = state, 1:2) %>% 
+  mutate(region = 2)
+
+# Add missing Missouri
+region_2[6, 2] <- "Missouri"
+
+region_3 <- regions_tbl[18:27,] %>%  
+  gather(key = region, value = state, 1:2) %>% 
+  slice(c(-10, -15, -16)) %>%
+  mutate(region = 3)
+
+region_4 <- 
+  regions_tbl[30:37,] %>%  
+  gather(key = region, value = state, 1:2) %>% 
+  slice(1:13) %>%
+  mutate(region = 4)
+
+regions <- rbind(region_1, region_2, region_3, region_4) %>%
+  mutate(state = str_to_title(state),
+         region = as.character(region))
+
+median_incomes <- clean_years %>%
+  group_by(region) %>%
+  summarise(zinc2_median = median(zinc2, na.rm = T))
+
+map_df <- map_data("state") %>%
+  mutate(region = str_to_title(region)) %>%
+  rename(state = region)  %>%
+  # Join with census region information
+  left_join(regions) %>%
+  left_join(median_incomes)
+
+usa_map <- map_df %>% 
+  ggplot() +  
+  geom_polygon(aes(x=long, 
+                   y = lat, 
+                   fill = zinc2_median, 
+                   group = group)) + 
+  coord_fixed(1.3) +
+  theme_minimal() +
+  scale_x_continuous(name = NULL,
+                     labels = NULL,
+                     breaks = NULL) +
+  scale_y_continuous(name = NULL,
+                     labels = NULL,
+                     breaks = NULL) +
+  labs(fill = "Median income") +
+  scale_fill_gradient2(midpoint = 25000)
+
+```
+
+```{r, echo=F}
+usa_map
+```
+
+To gauge income levels in different areas we used the four census regions in the dataset and joined them with a scraped table from mapsoftheworld.com.
+Using the table we could connect census regions to state names and thus colour a map of the continental US with census region incomes.
+We used median incomes instead of mean incomes to avoid skewing the results with outliers.
+
+The household median monthly incomes vary between different census regions from 32200 dollars in the South to nearly 25% higher, 40000, in the West.
+The Northeast trails closely behind the West with 37267.50 dollars, and the Midwest is closer to the South with a median income of 35000.
+
+While the dataset doesn't have more granular location data, even these four regions are very distinct from each other.
+The large high-income cities in the Northeast and the West likely contribute to the higher respective median incomes, while more rural Midwest and South are poorer.

### 2016-08-01 code/LambGraphs.R
lamb's graph

@@ -1,0 +1,79 @@
+
+
+##### BRANCH NAMES: LambTesting
+
+cleanyears <- read_csv("clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+  select(year, zsmhc, region, fmtburden) %>% 
+  filter(fmtburden != ".") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+
+
+
+##### Cleaning data for 
+cleanyearsmonthly <- read_csv("clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+  select(zinc2, year, region, fmtburden)%>% 
+  filter(fmtburden != ".") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+
+
+regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
+
+
+#### Monthly Housing Costs
+
+####overall graph
+
+cleanyears %>%  
+
+  group_by_("year", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
+  
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs", x = "Year", color = "Burden") 
+
+
+### faceted by region
+  cleanyears %>%  
+  
+  group_by_("year","region", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
+
+ 
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs", x = "Year", color = "Burden") 
+
+
+###### Household Income
+  ####overall graph
+cleanyearsmonthly %>%  
+    group_by_("year", "fmtburden") %>%  
[…]

### 2016-08-01 code/my_plots.Rmd
load plot codes

--- 2016-07-30 code/aplmed_region.Rmd
@@
 This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-IT is clear that the Northeast and the West are the most expensive regions.
+It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and WEst are by far the most productive (by GDP) regions in the United States.
+
 The median adjusted income increases steadily (roughly linearly) over time.
 There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.
+```{r}
+library(ggplot2)
+library(readr)
+library(stringr)
+library(dplyr)
+
+col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col)
+
+clean_years %>%
+  group_by(dataset_year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = dataset_year,   
+                  y = new_y,   
+                  colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "ABLmed by region")
+```
+
+The graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.

### 2016-08-01 code/ablmed_region_plot.R

@@ -1,0 +1,18 @@
+library(ggplot2)
+library(readr)
+library(stringr)
+library(dplyr)
+#col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv") #col_types = col)
+
+clean_years %>%
+  group_by(dataset_year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = dataset_year,   
+             y = new_y,   
+             colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "ABLmed by region")

### 2016-08-01 code/aplmed_region.Rmd
edit rmd
update plot codes

@@
 clean_years %>%
-  group_by(dataset_year, region) %>%  
+  group_by(year, region) %>%  
   # Take the median instead of mean to avoid extreme observations skewing the 
@@
   summarise(new_y = median(aplmed, na.rm = T)) %>%
-  ggplot(aes(x = dataset_year,   
+  ggplot(aes(x = year,   
                   y = new_y,   
@@
 This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-IT is clear that the Northeast and the West are the most expensive regions.
+It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and WEst are by far the most productive (by GDP) regions in the United States.
+
 The median adjusted income increases steadily (roughly linearly) over time.
 There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.
+```{r}
+library(ggplot2)
+library(readr)
+library(stringr)
+library(dplyr)
+
+col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col)
+
+clean_years %>%
+  group_by(dataset_year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = dataset_year,   
+                  y = new_y,   
+                  colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "ABLmed by region")
+```

### 2016-08-01 code/ablmed_region_plot.R

@@ -7,11 +7,11 @@ library(dplyr)
 clean_years <- read_csv("../data/clean_years.csv") #col_types = col)
 
 clean_years %>%
-  group_by(dataset_year, region) %>%  
+  group_by(year, region) %>%  
   # Take the median instead of mean to avoid extreme observations skewing the 
   # result.
   summarise(new_y = median(ablmed, na.rm = T)) %>%
-  ggplot(aes(x = dataset_year,   
+  ggplot(aes(x = year,   
              y = new_y,   
              colour = region)) + geom_line() +
   labs(x = "year", y = "Median Adjusted Income (per persons)", 

### 2016-08-02 code/William_final_section.Rmd

---
title: "Final Paper"
author: "William Paul"
date: "July 31, 2016"
output: html_document
---

```{r setup, include=FALSE, echo = F, message=F,error=F}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(readr)
library(stringr)
```

```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE}
clean_sample <- read_csv("../data/clean_years.csv")
clean_sample %>% filter(fmtstructuretype != ".") %>% 
      # Use NSE
      group_by_("year", "fmtstructuretype") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(x = "year",   
                        y = "new_y",   
                        colour = "fmtstructuretype")) +  
      geom_line() +
  labs(title="Income compared to area's median grouped by house type",
       colour = "Structure Type",
       y="Income percentage of area median income",
       x = "Year")
```
This plot charts a time series showing how the typical income for various types of households such as single family or apartment buildings changed over the years. One major trend is that typical income of all housing types except single family homes compared to their area has fallen, accelerating after 2007. However, incomes of single family homes, barring the housing crash in 2007, have been centered on the area's median income. There are a couple of possible explanations for this disparity. The first is that, with the growth of wages, more people could afford single family homes and thus leave the buildings with higher numbers of units. As wage grew, the area's median income also grew, and so, combined with higher waged workers moving into single family houses, single family homes remained constant while all other types shrunk proportionally. This would also explain the acceleration after the housing crisis, as single family homes became much cheaper temporarily and thus allow more workers to obtain single family housing. The second explanation is that as time went on, supply of housing increased, driving prices down and allowing lower waged laborers to obtain homes. Due to the purchasing power of higher wage laborers however, single family houses, even with their falling prices, were more difficult to obtain as time went on, and thus lower wage workers had to settle for other types of housing. 
```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
clean_sample %>% filter(assisted == 0) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(y = "new_y",   
                        x = "fmtcostmedrelamicat",
                        fill = "fmtcostmedrelamicat")) +  
      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost without assistance",
       fill = "Cost as Percentage of Area Median Income",
       y="Burden",
       x = "") + 
  scale_y_continuous(limits=c(0,1)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_flip()
clean_sample %>% filter(assisted == 1) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(y = "new_y",   
                        x = "fmtcostmedrelamicat",
                        fill = "fmtcostmedrelamicat")) +  
      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost with assistance",
       fill = "Cost as Percentage of Area Median Income",
       y="Burden",
       x = "") + 
  scale_y_continuous(limits=c(0,1)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_flip()
```
The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs. At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance. This clearly makes sense, as those who need assistance are typically people who've lost their job or had some monetary crisis that would reduce their total salary. However, although assistance has around the same effect for most cost brackets, the households that have excessively high costs for their area have vastly higher burdens with assistance than all other brackets. One explanation for this is that those who purchase expensive houses in their area typically have higher salaries, and if they lose those salaries and require assistance they have to support the cost of the house. Also, single family homeowners have significantly less mobility so if they have a financial crisis they cannot easily transition to cheaper housing as opposing to those with rental properties typical of a larger number of units.

### 2016-08-02 code/my_plots.Rmd

@@
   summarise(new_y = median(ablmed, na.rm = T)) %>%
-  ggplot(aes(x = dataset_year,   
+  ggplot(aes(x = year,   
                   y = new_y,   

### 2016-08-02 code/Paper.Rmd
current changes

--- 2016-08-01 code/nikolais_plot.Rmd
@@
 ---
+title: "Paper TITLE"
 output: html_document
 ---
-```{r, include=F}
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+
 library(ggplot2)
@@
 
-```{r, echo=F}
+```{r, echo = F}
 col_types <- "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
@@
          y = "Median burden")
+
 ```
@@
 
-```{r, include=F}
+```{r, include = F}
 url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
@@
 The large high-income cities in the Northeast and the West likely contribute to the higher respective median incomes, while more rural Midwest and South are poorer.
+
+```{r, include = FALSE}
+regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
+
+```
+
+```{r, include = FALSE}
+
+cleanyears <- read_csv("../data/clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+  select(zinc2, year, region, fmtburden)%>% 
+  filter(fmtburden != ".") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+
+###### Household Income
+  ####overall graph
+cleanyears %>%  
+    group_by_("year", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zinc2", ",na.rm=T)"))) %>%
+
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  labs(title = "Relationship between Household \nIncome and Burden", 
+       y = "Household Income ($)", x = "Year", color = "Burden")
+
+```
+
+```{r, include = FALSE}
+
+ ### faceted by region 
+  cleanyears %>%  
+  group_by_("year","region", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zinc2", ",na.rm=T)"))) %>%
+
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  facet_wrap(~region, labeller = as_labeller(regionNames)) +
+  labs(title = "Relationship between Household \nIncome and Burden", 
+       y = "Household Income ($)", x = "Year", color = "Burden")
+
+```
+
+```{r, include = FALSE}
+
+cleanyearsmonthly <- read_csv("../data/clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+  select(year, zsmhc, region, fmtburden) %>% 
+  filter(fmtburden != ".") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+
+####overall graph
+
+cleanyearsmonthly %>%
+
+  group_by_("year", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%
+
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden")
+
+```
+
+```{r, include = FALSE}
+
+### faceted by region
+  cleanyearsmonthly %>%
+
+  group_by_("year","region", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%
+
+  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  geom_line() + 
+  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden")
+
+```

### 2016-08-02 code/LambGraphs.R
current changes

@@ -2,6 +2,9 @@
 
 ##### BRANCH NAMES: LambTesting
 
+
+
+##### Cleaning data for monthly costs
 cleanyears <- read_csv("clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
   select(year, zsmhc, region, fmtburden) %>% 
   filter(fmtburden != ".") %>% 
@@ -9,7 +12,7 @@ cleanyears <- read_csv("clean_years.csv", col_types = "cicciiciiciiidicidicddddd
 
 
 
-##### Cleaning data for 
+##### Cleaning data for household income
 cleanyearsmonthly <- read_csv("clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
   select(zinc2, year, region, fmtburden)%>% 
   filter(fmtburden != ".") %>% 

### 2016-08-02 code/my_plots.Rmd
edit plots

@@
 It is clear that the Northeast and the West are the most expensive regions.
-This is makes sense given that the northeast and WEst are by far the most productive (by GDP) regions in the United States.
+This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
 
 The median adjusted income increases steadily (roughly linearly) over time.
-There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
+economy) while the other regions took some time to correct for the downturn in the financial market.
+
 ```{r}
@@
 
-The graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+
+It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
+This is not shown in the graphs, which indicate that the income levels are significantly different.
+However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
+
+The lag in the downward trend after 2008 still occurs in the plots.

### 2016-08-02 code/Paper.Rmd
recent change
chaaange

@@
 ```
+
+Summary/Intro
+
+We chose the dataset from the American Housing Survey, which derives sets of data from 1985 and on.
+The data set includes categorizing each housing unit by factors such as affordability, income through Fair Market Rent and the Adjusted Median Income.
+Taking into account variables such as Burden (housing cost divided by its monthly income), certain households with a zero or negative income are input as -1 within the data.
+The data source was recorded from the American Housing Survey.
+Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+
+We chose to analyze the data and ultimately view determines whether housing affordability has increased or decreased over the last few decades.
+Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
+The United State’s growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
+Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
+Between all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the housing market affordability status.
+
+Cleaning data
+
+The "combined_ years" data was read into a dataframe using the "read_ csv" function.
+Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), and the nonsensical integer date (negative numbers) were converted to NAs in the dataset.
+Then, the combined years dataset was cleaned using the piping methods in dplyr.
+All the column names were set to lowercase, and all the quotations were removed from character values.
+Then, an additional column “ownrent” was added which labeled levels 1 and 2 to own and rent as factors. Then, the data was written to a clean_years csv file.
 
@@
 
-```{r, include = FALSE}
+```{r, echo = FALSE}
 regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
@@
 
-```{r, include = FALSE}
+```{r, echo = FALSE}
 
@@
 
-```{r, include = FALSE}
+```{r, echo= FALSE}
 
@@
 ```
+
+```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE}
+
+clean_sample <- read_csv("../data/clean_years.csv")
+clean_sample %>% filter(fmtstructuretype != ".") %>% 
+      # Use NSE
+      group_by_("year", "fmtstructuretype") %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
+      # Use NSE
+      ggplot(aes_string(x = "year",   
+                        y = "new_y",   
+                        colour = "fmtstructuretype")) +  
+      geom_line() +
+  labs(title="Income compared to area's median grouped by house type",
+       colour = "Structure Type",
+       y="Income percentage of area median income",
+       x = "Year")
+
+```
+
+This plot charts a time series showing how the typical income for various types of households such as single family or apartment buildings changed over the years.
+One major trend is that typical income of all housing types except single family homes compared to their area has fallen, accelerating after 2007.
+However, incomes of single family homes, barring the housing crash in 2007, have been centered on the area's median income.
+There are a couple of possible explanations for this disparity.
+The first is that, with the growth of wages, more people could afford single family homes and thus leave the buildings with higher numbers of units.
+As wage grew, the area's median income also grew, and so, combined with higher waged workers moving into single family houses, single family homes remained constant while all other types shrunk proportionally.
+This would also explain the acceleration after the housing crisis, as single family homes became much cheaper temporarily and thus allow more workers to obtain single family housing.
+The second explanation is that as time went on, supply of housing increased, driving prices down and allowing lower waged laborers to obtain homes.
+Due to the purchasing power of higher wage laborers however, single family houses, even with their falling prices, were more difficult to obtain as time went on, and thus lower wage workers had to settle for other types of housing.
+
+```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
+
+clean_sample %>% filter(assisted == 0) %>% 
+      # Use NSE
+      group_by_("assisted", "fmtcostmedrelamicat") %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
+      # Use NSE
+      ggplot(aes_string(y = "new_y",   
+                        x = "fmtcostmedrelamicat",
+                        fill = "fmtcostmedrelamicat")) +  
+      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost without assistance",
+       fill = "Cost as Percentage of Area Median Income",
+       y="Burden",
+       x = "") + 
+  scale_y_continuous(limits=c(0,1)) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  coord_flip()
+clean_sample %>% filter(assisted == 1) %>% 
+      # Use NSE
+      group_by_("assisted", "fmtcostmedrelamicat") %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
+      # Use NSE
+      ggplot(aes_string(y = "new_y",   
+                        x = "fmtcostmedrelamicat",
+                        fill = "fmtcostmedrelamicat")) +  
+      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost with assistance",
+       fill = "Cost as Percentage of Area Median Income",
+       y="Burden",
+       x = "") + 
+  scale_y_continuous(limits=c(0,1)) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  coord_flip()
+
+```
+
+The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
+At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance.
+This clearly makes sense, as those who need assistance are typically people who've lost their job or had some monetary crisis that would reduce their total salary.
+However, although assistance has around the same effect for most cost brackets, the households that have excessively high costs for their area have vastly higher burdens with assistance than all other brackets.
+One explanation for this is that those who purchase expensive houses in their area typically have higher salaries, and if they lose those salaries and require assistance they have to support the cost of the house.
+Also, single family homeowners have significantly less mobility so if they have a financial crisis they cannot easily transition to cheaper housing as opposing to those with rental properties typical of a larger number of units.
+
+```{r}
+
+```

### 2016-08-02 code/Paper.Rmd
paper changes

@@
 
-```{r}
+```{r, echo = FALSE}
+
+col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "APLmed by region")
 
 ```
+
+This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and WEst are by far the most productive (by GDP) regions in the United States.
+
+The median adjusted income increases steadily (roughly linearly) over time.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.
+
+```{r, echo = F}
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "ABLmed by region")
+
+```
+
+The graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.

### 2016-08-03 code/Papeeeer.Rmd
pap

--- 2016-08-02 code/Paper.Rmd
@@
 title: "Paper TITLE"
-output: html_document
+output: pdf_document
 ---
@@
 Cleaning data
+
+Cleaning data
+Initially, the dataset is split into fifteen datasets, one for each odd year between 1985 and 2013.
+Due to differences in how the survey was conducted over the years, each dataset has varying column names and positions, and, for surveys performed earlier than 2003, missing columns.
+To standardize the data, we used the format of the more modern datasets, i.e the datasets from 2003 to 2013.
+Thus, we begin by appending a column onto each dataset to mark which year that data was gathered, remove an extraneous column from the 2009 dataset, and then simply combine the datasets.
+For the older years, the datasets must again be split into two groups as some columns are named differently and do not fit with our standard.
+Once we rename the columns, we add the missing columns that dictate the raw, adjusted by growth, and adjusted by people income limits for each household.
+These income limits specifically determine what the extremely low income bracket is for the area each household is in.
+For each list of datasets, we reorder the columns to exactly match that of our base, the datasets gathered from 2003 onward.
+Finally we combine every list of datasets into a single dataset, and then store it in a comma separated file called combined_years.
 
@@
 
+As pictured in the graph, the relationship between household income and burden is seen as a positive correlation.
+From the graph, you can see that higher income household have less burden to deal with, which makes sense because those with more money are able to pay off monthly costs more easily.
+Those with 50% more burden are near the poverty line.
+From this graph, you can see a spike in households that have burden.
+Households with higher incomes started to pay more burdens, which can be attributed to 2007 financial housing market crisis.
+
 ```{r, echo = FALSE}
@@
 ```
+
+From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
 
@@
 
+From this graph, you can take a more indepth look at the relationship between housing costs and burden.
+It's interesting to see that all three burden lines are close to one another, which illustrates the principle that those with more money aren't buying more expensive houses.
+From the previous grpah, we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
+Each level of burden has similar housing costs, therefore reinforcing the idea that people are buying similar priced housing among the lower and middle classes.
+Again, a spike around 2010 is prevalent, which can be caused by the 2007 housing crisis.
+
 ```{r, echo= FALSE}
@@
 The graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+
+Conclusion:
+
+Blah blah blah blah blah

### 2016-08-03 code/parse_data.R
Added X axis to shiny and shuffling to clean years

@@ -27,7 +27,7 @@ remove(combined_years)
 # Make a sample for faster developing etc.
 set.seed(1337)  
 clean_sample <- sample_n(clean_years, 1000)
-
+clean_years = clean_years[sample(nrow(clean_years)),]
 # Write csv files
 write_csv(clean_years, path = "../data/clean_years.csv")
 write_csv(clean_sample, path = "../data/clean_sample.csv")

### 2016-08-03 code/server.R

@@ -10,14 +10,14 @@ shinyServer(function(input, output) {
   output$plot <- renderPlot({ 
      clean_sample %>%  
       # Use NSE
-      group_by_("year", input$sel_x) %>%  
+      group_by_(input$sel_x, input$sel_color) %>%  
       # Take the median instead of mean to avoid extreme observations skewing the 
       # result.
       summarise_(new_y= as.formula(paste0("~ median(",input$sel_y, ",na.rm=T)"))) %>%  
       # Use NSE
-      ggplot(aes_string(x = "year",   
+      ggplot(aes_string(x = input$sel_x,   
                         y = "new_y",   
-                        colour = input$sel_x)) +  
+                        colour = input$sel_color)) +  
       geom_line()
 
   })

### 2016-08-04 code/parse_data.R
Add new plot, change '.' to NA

@@ -6,7 +6,7 @@ library(tidyr)
 combined_years <- read_csv("../data/combined_years.csv",
                            # Add col_types manually to avoid parsing failures
                            col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd",
-                           na = c("","NA", "-9", "-6", "-7"))
+                           na = c("","NA", "-9", "-6", "-7", "."))
 
 NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")

### 2016-08-04 code/parse_data.R
Substitute '.'s with NAs, working

@@ -6,7 +6,7 @@ library(tidyr)
 combined_years <- read_csv("../data/combined_years.csv",
                            # Add col_types manually to avoid parsing failures
                            col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd",
-                           na = c("","NA", "-9", "-6", "-7", "."))
+                           na = c("","NA", "-9", "-6", "-7", "'.'"))
 
 NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")

### 2016-08-04 code/Paper.Rmd
changed paper name

@@
 It is clear that the Northeast and the West are the most expensive regions.
-This is makes sense given that the northeast and WEst are by far the most productive (by GDP) regions in the United States.
+This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
 
 The median adjusted income increases steadily (roughly linearly) over time.
-There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
+economy) while the other regions took some time to correct for the downturn in the financial market.
 
@@
 
-The graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+
+It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
+This is not shown in the graphs, which indicate that the income levels are significantly different.
+However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
+
+The lag in the downward trend after 2008 still occurs in the plots.
+
+Conclusion:
+
+Blah blah blah blah blah

### 2016-08-04 code/Paper.Rmd

@@
   labs(fill = "Median income") +
-  scale_fill_gradient2(midpoint = 25000)
+  scale_fill_distiller(palette = "Blues",
+                       direction = 1)
 

### 2016-08-04 code/Paper.Rmd
Add plot for poverty groups over time

@@
 The large high-income cities in the Northeast and the West likely contribute to the higher respective median incomes, while more rural Midwest and South are poorer.
+
+```{r, echo = FALSE}
+counts <- clean_years %>% 
+  # Group and summarise the amount of people in each poverty group each year
+  group_by(fmtincrelpovcat, year) %>%
+  summarise(count = n()) %>%
+  # Clean NAs
+  na.omit() %>%
+  # Group again by year to calculate relative percentages per year
+  group_by(year) %>% 
+  mutate(percent = count / sum(count),  
+         # Clean the labels
+         fmtincrelpovcat = factor(fmtincrelpovcat, 
+                                  levels = c("1 LTE Poverty", 
+                                             "2 100-150% Poverty", 
+                                             "3 150-200% Poverty", 
+                                             "4 200%+ Poverty"),  
+                                  labels = c("Less than poverty income",   
+                                             "100-150% of poverty income",   
+                                             "150-200% of poverty income",    
+                                             "> 200% of poverty income"),
+                                  ordered = T))
+
+ggplot(counts, aes(x = year, 
+                   y = percent,
+                   fill = fmtincrelpovcat)) + 
+  geom_area(alpha = 0.5) +
+  # Use a nice colour palette
+  scale_fill_brewer(palette = "RdBu", 
+                    # Reverse the order of legend items to match the
+                    # stacking order
+                    guide = guide_legend(reverse=TRUE)) +
+  labs(title = "Amount of households relative to poverty line",
+       x = "Year",
+       y = "Precent of households",
+       fill = "Household income
+relative to poverty income") 
+```
 

### 2016-08-04 code/Paper.Rmd
minor changes to Sri Plots

--- 2016-08-04 code/Paper.Rmd
@@
                   colour = region)) + geom_line() + 
-  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+  scale_fill_manual(name = "Region", labels = regionNames)
+  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
        title = "ABLmed by region")

### 2016-08-04 code/Paper.Rmd
Changed histogram colors and overlayed both histograms together

@@
 ```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
-
-clean_sample %>% filter(assisted == 0) %>% 
+assisted = clean_sample %>% filter(!is.na(assisted)) %>% 
       # Use NSE
@@
       summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
-      # Use NSE
-      ggplot(aes_string(y = "new_y",   
-                        x = "fmtcostmedrelamicat",
-                        fill = "fmtcostmedrelamicat")) +  
-      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost without assistance",
-       fill = "Cost as Percentage of Area Median Income",
+      ungroup %>% 
+      mutate(assisted = as.factor(assisted))
+ggplot(assisted, aes(y = new_y,   
+                  x = fmtcostmedrelamicat,
+                  fill = assisted, group = assisted)) +  
+  geom_bar(
+           stat="identity") + 
+  labs(title="Burden based on Median housing cost with assistance",
        y="Burden",
-       x = "") + 
+       x = "",
+       fill="Assistance") + 
   scale_y_continuous(limits=c(0,1)) +
-  guides(fill = guide_legend(reverse = TRUE)) +
+  scale_fill_manual("Assistance",labels=c("Without Assistance", "With Assistance"), values=c("black", "white")) +
   coord_flip()
-clean_sample %>% filter(assisted == 1) %>% 
-      # Use NSE
-      group_by_("assisted", "fmtcostmedrelamicat") %>%  
-      # Take the median instead of mean to avoid extreme observations skewing the 
-      # result.
-      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
-      # Use NSE
-      ggplot(aes_string(y = "new_y",   
-                        x = "fmtcostmedrelamicat",
-                        fill = "fmtcostmedrelamicat")) +  
-      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost with assistance",
-       fill = "Cost as Percentage of Area Median Income",
-       y="Burden",
-       x = "") + 
-  scale_y_continuous(limits=c(0,1)) +
-  guides(fill = guide_legend(reverse = TRUE)) +
-  coord_flip()
-
 ```

### 2016-08-04 code/Paper.Rmd
changed more of Sri's stuff

--- 2016-08-04 code/Paper.Rmd
@@
                   colour = region)) + geom_line() +
-  labs(x = "year", y = "Median Adjusted Income (per persons)", 
-       title = "APLmed by region")
+  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
+       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  scale_color_discrete(name = "Region", labels = regionNames)
 
@@
                   colour = region)) + geom_line() +
-  scale_fill_manual(name = "Region", labels = regionNames)
   labs(x = "Year", y = "Median Adjusted Income (per persons)", 
-       title = "ABLmed by region")
+       title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
+  scale_color_discrete(name = "Region", labels = regionNames)
 
@@
 
-It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+It is interesting to note that the disparity between the incomes increases between the Northeast/Western regions and the Midwest/Southern Regions.
 This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.

### 2016-08-04 code/Paper.Rmd
Fine-tune map colours

@@
 library(maps)
+library(ggthemes)
 ```
@@
   labs(fill = "Median income") +
-  scale_fill_distiller(palette = "Blues",
-                       direction = 1)
-
+  scale_fill_distiller(limits = c(31000, 41200), 
+                                 breaks = sort(unique(map_df$zinc2_median),
+                                               decreasing = T),
+                                 palette = "Blues", 
+                                 direction = 1,
+                                 guide = "legend")
 ```

### 2016-08-05 code/Paper.Rmd
Add themes, tune colours

@@
          x = "Year",
-         y = "Median burden")
+         y = "Median burden") +
+    theme_tufte()
 
@@
   labs(fill = "Median income") +
-  scale_fill_distiller(limits = c(31000, 41200), 
+  scale_fill_distiller(limits = c(25000, 42000), 
                                  breaks = sort(unique(map_df$zinc2_median),
                                                decreasing = T),
-                                 palette = "Blues", 
+                                 palette = "Greens", 
                                  direction = 1,
@@
                    fill = fmtincrelpovcat)) + 
-  geom_area(alpha = 0.5) +
+  geom_area(alpha = 0.9) +
   # Use a nice colour palette
-  scale_fill_brewer(palette = "RdBu", 
+  scale_fill_brewer(palette = "RdYlGn", 
                     # Reverse the order of legend items to match the
@@
        fill = "Household income
-relative to poverty income") 
+relative to poverty income") +
+  theme_tufte()
 ```
@@
   labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income ($)", x = "Year", color = "Burden")
+       y = "Household Income ($)", x = "Year", color = "Burden") +
+  theme_tufte()
 
@@
   labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
-       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden")
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  theme_tufte()
 
@@
        y="Income percentage of area median income",
-       x = "Year")
+       x = "Year") +
+  theme_tufte()
 
@@
   scale_fill_manual("Assistance",labels=c("Without Assistance", "With Assistance"), values=c("black", "white")) +
-  coord_flip()
+  coord_flip() +
+  theme_tufte()
 ```
@@
        title = "Median Income Adjusted \nfor the Number of People by Region") + 
-  scale_color_discrete(name = "Region", labels = regionNames)
+  scale_color_discrete(name = "Region", labels = regionNames) +
+  theme_tufte()
 
@@
        title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
-  scale_color_discrete(name = "Region", labels = regionNames)
+  scale_color_discrete(name = "Region", labels = regionNames) +
+  theme_tufte()
 

### 2016-08-05 code/Paper.Rmd
Modify barchart colour scheme

@@
   scale_y_continuous(limits=c(0,1)) +
-  scale_fill_manual("Assistance",labels=c("Without Assistance", "With Assistance"), values=c("black", "white")) +
+  scale_fill_brewer("Assistance", 
+                    labels = c("Without Assistance", 
+                               "With Assistance"),
+                    palette = "Set2")  +
   coord_flip() +

### 2016-08-05 code/Paper.Rmd
Modify alpha in map

@@
                    fill = fmtincrelpovcat)) + 
-  geom_area(alpha = 0.9) +
+  geom_area(alpha = 0.8) +
   # Use a nice colour palette

### 2016-08-05 code/Paper.Rmd
Change colour themes

@@
                       colour = per)) +  
-    geom_line() +
+    geom_line(size = 1) +
     labs(colour = "Household size",
@@
          y = "Median burden") +
+    scale_colour_brewer(palette = "Dark2") +
     theme_tufte()
@@
 relative to poverty income") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line() + 
+  geom_line(size = 1) + 
   labs(title = "Relationship between Household \nIncome and Burden", 
        y = "Household Income ($)", x = "Year", color = "Burden") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line() + 
+  geom_line(size = 1) + 
   facet_wrap(~region, labeller = as_labeller(regionNames)) +
@@
        y = "Household Income ($)", x = "Year", color = "Burden") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line() + 
+  geom_line(size = 1) + 
   labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
        y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line() + 
+  geom_line(size = 1) + 
   facet_wrap(~region, labeller = as_labeller(regionNames)) + 
@@
        y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
                         colour = "fmtstructuretype")) +  
-      geom_line() +
+  geom_line(size = 1) + 
   labs(title="Income compared to area's median grouped by house type",
@@
        x = "Year") +
+  scale_colour_brewer(palette = "Dark2") +
   theme_tufte()
@@
                   y = new_y,   
-                  colour = region)) + geom_line() +
+                  colour = region)) + 
+  geom_line(size = 1) + 
   labs(x = "Year", y = "Median Adjusted Income (per persons)", 
        title = "Median Income Adjusted \nfor the Number of People by Region") + 
-  scale_color_discrete(name = "Region", labels = regionNames) +
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
   theme_tufte()
@@
                   y = new_y,   
-                  colour = region)) + geom_line() +
+                  colour = region)) + 
+  geom_line(size = 1) +
   labs(x = "Year", y = "Median Adjusted Income (per persons)", 
        title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
-  scale_color_discrete(name = "Region", labels = regionNames) +
+  scale_colour_brewer(palette = "Dark2",
+                      name = "Region", 
+                      labels = regionNames) +
   theme_tufte()

### 2016-08-05 code/Paper.Rmd
minor grammer

--- 2016-08-05 code/Paper.Rmd
@@
 library(maps)
-library(ggthemes)
 ```
@@
          x = "Year",
-         y = "Median burden") +
-    theme_tufte()
+         y = "Median burden")
 
@@
   labs(fill = "Median income") +
-  scale_fill_distiller(limits = c(25000, 42000), 
-                                 breaks = sort(unique(map_df$zinc2_median),
-                                               decreasing = T),
-                                 palette = "Greens", 
-                                 direction = 1,
-                                 guide = "legend")
+  scale_fill_distiller(palette = "Blues",
+                       direction = 1)
+
 ```
@@
                    fill = fmtincrelpovcat)) + 
-  geom_area(alpha = 0.8) +
+  geom_area(alpha = 0.5) +
   # Use a nice colour palette
-  scale_fill_brewer(palette = "RdYlGn", 
+  scale_fill_brewer(palette = "RdBu", 
                     # Reverse the order of legend items to match the
@@
        fill = "Household income
-relative to poverty income") +
-  theme_tufte()
+relative to poverty income") 
 ```
@@
   labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income ($)", x = "Year", color = "Burden") +
-  theme_tufte()
+       y = "Household Income ($)", x = "Year", color = "Burden")
 
@@
   labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
-       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
-  theme_tufte()
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden")
 
@@
        y="Income percentage of area median income",
-       x = "Year") +
-  theme_tufte()
+       x = "Year")
 
@@
   scale_y_continuous(limits=c(0,1)) +
-  scale_fill_brewer("Assistance", 
-                    labels = c("Without Assistance", 
-                               "With Assistance"),
-                    palette = "Set2")  +
-  coord_flip() +
-  theme_tufte()
+  scale_fill_manual("Assistance",labels=c("Without Assistance", "With Assistance"), values=c("black", "white")) +
+  coord_flip()
 ```
@@
        title = "Median Income Adjusted \nfor the Number of People by Region") + 
-  scale_color_discrete(name = "Region", labels = regionNames) +
-  theme_tufte()
+  scale_color_discrete(name = "Region", labels = regionNames)
 
@@
        title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
-  scale_color_discrete(name = "Region", labels = regionNames) +
-  theme_tufte()
+  scale_color_discrete(name = "Region", labels = regionNames)
 

### 2016-08-06 code/Paper.Rmd
minor grammar changes

--- 2016-08-05 code/Paper.Rmd
@@
   geom_line(size = 1) + 
-  labs(title = "Relationship between Household \nIncome and Burden", 
+  labs(title = "Figure : Relationship between Household \nIncome and Burden", 
        y = "Household Income ($)", x = "Year", color = "Burden") +

### 2016-08-06 paper/Paper.Rmd
Move files to right dirs

--- 2016-08-06 code/Paper.Rmd
@@
   coord_fixed(1.3) +
-  theme_minimal() +
+  theme_tufte() +
   scale_x_continuous(name = NULL,
@@
 ```
+
+This plot shows the amount of households in a given income group in a given year.
+Each household has been calculated a poverty treshold according to the Census Bureau’s official thresholds, based on family size and number of children.
+
+As can be seen from the plot, the amount of people in different income groups stayed relatively constant between 1985 and 2007.
+The majority of people earned over twice the poverty income, while little under a fifth of the population was in poverty.
+This stable relationship broke noticeably in the financial crisis of 2007 and the following recession, when the amount of people in poverty started a quite steep and steady climb upwards.
+
+Interestingly, the income brackets of 100%-200% stayed almost exactly the same size as before, while the increase in poverty was in effect solely the result of the dwindling of the highest (over 200% of poverty income) income bracket.
+A reasonable explanation would be the effect of increasing unemployment, with previously employed well-off people dropping to unemployment and thus small or no income.
+The reality can be more nuanced, with people dropping to the middle income groups and a simultaneous equally large movement from these brackets to poverty, but the big picture of large impacts in poverty following the financial crisis stays the same.
 

### 2016-08-06 code/ablmed_region_plot.R
Remove unnecessary files

@@ -1,18 +1,0 @@
-library(ggplot2)
-library(readr)
-library(stringr)
-library(dplyr)
-#col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
-
-clean_years <- read_csv("../data/clean_years.csv") #col_types = col)
-
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(ablmed, na.rm = T)) %>%
-  ggplot(aes(x = year,   
-             y = new_y,   
-             colour = region)) + geom_line() +
-  labs(x = "year", y = "Median Adjusted Income (per persons)", 
-       title = "ABLmed by region")

### 2016-08-06 code/Paper.Rmd
minor grammer

--- 2016-08-05 code/Paper.Rmd
@@
 library(purrr)
+library(scales)
 library(maps)
@@
 
-cleanyears <- read_csv("../data/clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+cleanyears <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
   select(zinc2, year, region, fmtburden)%>% 
@@
   geom_line(size = 1) + 
-  labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income ($)", x = "Year", color = "Burden") +
+  labs(title = "Figure : Relationship between Household \nIncome and Burden", 
+       y = "Household Income", x = "Year", color = "Burden") +
+  scale_y_continuous(labels = unit_format(unit = "K")) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-cleanyearsmonthly <- read_csv("../data/clean_years.csv", col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd") %>% 
+cleanyearsmonthly <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
   select(year, zsmhc, region, fmtburden) %>% 
@@
 
-col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
-
 clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)

### 2016-08-07 clean_data/CleanData.R
Merge combining and cleaning data to a single file
The paths to the files have also been modified in the shiny app
and the gitignore file. Codebook.txt added to describe the data.

@@ -1,12 +1,117 @@
+# Running this code may take up to ten minutes because of the large size
+# of the datasets. Running from terminal instead of RStudio
+# ("Rscript CleanData.R") or on a powerful server is recommended.
+
 library(readr)
 library(dplyr)
 library(stringr)
 library(tidyr)
 
-combined_years <- read_csv("../data/combined_years.csv",
-                           # Add col_types manually to avoid parsing failures
-                           col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd",
-                           na = c("","NA", "-9", "-6", "-7", "'.'"))
+MutateYears <- function(data, year) {
+  # Simply marks the dataset with the year to categorize the original data.
+  mutate(data, 
+         YEAR = year)
+}
+
+UpperCaseColNames <- function(frame) {
+  # Years vary over the years, so this converts the column 
+  # names to the same style as that of the latest dataset.
+  setNames(frame, 
+           (toupper(names(frame)) %>% 
+              str_replace("^COSTMED(.+)",
+                          "COSTMed\\1")))
+}
+
+AddMissingCols <- function(frame) {
+  # Datasets before 2003 are missing the three columns below, 
+  # so this function adds them to unify all of the datasets
+  mutate(frame, 
+         L30 = 0.6 * L50, 
+         GL30 = 0.6 * GL50, 
+         ABL30 = 0.6 * ABL50)
+}
+
+RenameCols <- function(frame) {
+  # Naming convention changed after the 1995 dataset, 
+  # so we match the column names.
+  rename(frame, 
+         "AGE1" = `AGE`, 
+         "FMTMETRO3" = `FMTMETRO`, 
+         "STATUS" = `ISTATUS`, 
+         "METRO3" = `METRO`)
+}
+
+base_files <- paste("../raw_data/hads",          # Loading the most recent datasets 
+                c(seq(2003, 2013, by = 2)), # first to give reference for other
+                ".bz2",                     # two sets.
+                sep = "") %>%  
+  lapply(read_csv) %>% 
+  lapply(data.frame,  # Have to convert to data frames as lapply seems 
+         stringsAsFactors = F) #to recognize csv output as named lists.
+
+base_files <- mapply(MutateYears, # Need to distinguish different years
+                 base_files,     #for datasets, so add column to do just that.
[…]

### 2016-08-07 paper/Paper.Rmd

@@
   labs(title = "Figure : Relationship between Household \nIncome and Burden", 
-       y = "Household Income ($)", x = "Year", color = "Burden") +
+       y = "Household Income", x = "Year", color = "Burden") +
+  scale_y_continuous(labels = unit_format(unit = "K")) +
   scale_colour_brewer(palette = "Dark2") +

### 2016-08-07 clean_data/CleanData.R
Re-add sample generation to CleanData.R

@@ -136,9 +136,12 @@ clean_years <- combined_years %>%
 # Free some memory
 remove(combined_years)
 
-# Randomize rows to avoid column type detection problems when reading data
 set.seed(1337)  
+# Make a sample for demonstration purposes
+clean_sample <- sample_n(clean_years, 10000)
+# Randomize rows to avoid column type detection problems when reading data
 clean_years <- sample_n(clean_years, nrow(clean_years))
 
 # Save the results
+write_csv(clean_sample, path = "clean_sample.csv")
 write_csv(clean_years, path = "clean_years.csv")

### 2016-08-07 code/Paper.Rmd
paragraph changes

@@
 
-The "combined_ years" data was read into a dataframe using the "read_ csv" function.
-Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), and the nonsensical integer date (negative numbers) were converted to NAs in the dataset.
+  The "combined-years" data was read into a dataframe using the read.csv function.
+Within the read.csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
+Also, the nonsensical integer date (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
 Then, the combined years dataset was cleaned using the piping methods in dplyr.
-All the column names were set to lowercase, and all the quotations were removed from character values.
-Then, an additional column “ownrent” was added which labeled levels 1 and 2 to own and rent as factors. Then, the data was written to a clean_years csv file.
+The data was rather messy, with unecessary quotaions around character values and inconsistent letter cases in the column names.
+This was rectified by setting all the column names to lowercase and removing all the quotations from character values.
+One the main premises of this paper is comparing the burden of household that are owning and renting.
+Since there was no explicit labeling of data corresponding to owning and renting households, an  additional column "ownrent" was created which labeled levels 1 and 2 to own and rent as factors.
+Finally, the data was written to a clean_years csv file.
 
@@
 This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-It is clear that the Northeast and the West are the most expensive regions.
-This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+APLmed is the adjusted median household income per person.
+This was precaculated using Housing and Urban Development (HUD) indexes.
+For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
+The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+
+It is clear that the Northeast and the West are the most expensive regions. This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
 
@@
   geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
+  labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)", 
        title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
@@
 The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+ABlmed was calculated using precaculated HUD indexes.
+Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
+Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
 
-It is interesting to note that the disparity between the incomes increases between the Northeast/Western regions and the Midwest/Southern Regions.
-This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+This is liekly because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
 However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
@@
 
-The lag in the downward trend after 2008 still occurs in the plots.
-
 Conclusion:

### 2016-08-07 code/Paper.Rmd
reorganized paper

@@
 
-Cleaning data
 Initially, the dataset is split into fifteen datasets, one for each odd year between 1985 and 2013.
@@
 
-  The "combined-years" data was read into a dataframe using the read.csv function.
-Within the read.csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
+The combined_years data was read into a dataframe using the read_csv function.
+Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
 Also, the nonsensical integer date (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
@@
 The large high-income cities in the Northeast and the West likely contribute to the higher respective median incomes, while more rural Midwest and South are poorer.
+
+```{r, echo = FALSE}
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
+       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte()
+
+```
+This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+APLmed is the adjusted median household income per person.
+This was precaculated using Housing and Urban Development (HUD) indexes.
+For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
+The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+
+It is clear that the Northeast and the West are the most expensive regions. This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+
+The median adjusted income increases steadily (roughly linearly) over time.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
+economy) while the other regions took some time to correct for the downturn in the financial market.
+
+```{r, echo = F}
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) +
+  labs(x = "Year", y = "Median Adjusted Income (per bedroom)", 
+       title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
+  scale_colour_brewer(palette = "Dark2",
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte()
+
+```
+
+The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+ABlmed was calculated using precaculated HUD indexes.
+Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
+Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
+
+It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+This is liekly because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
+This is not shown in the graphs, which indicate that the income levels are significantly different.
+However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
+
+The lag in the downward trend after 2008 still occurs in the plots.
 
@@
 
-```{r, echo = FALSE}
-
-clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
-
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(aplmed, na.rm = T)) %>%
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
-       title = "Median Income Adjusted \nfor the Number of People by Region") + 
-  scale_colour_brewer(palette = "Dark2", 
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_tufte()
-
-```
-
-This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-APLmed is the adjusted median household income per person.
-This was precaculated using Housing and Urban Development (HUD) indexes.
-For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
-The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
-
-It is clear that the Northeast and the West are the most expensive regions. This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
-
-The median adjusted income increases steadily (roughly linearly) over time.
-There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
-So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
-economy) while the other regions took some time to correct for the downturn in the financial market.
-
-```{r, echo = F}
-
-clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
-
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(ablmed, na.rm = T)) %>%
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)", 
-       title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
-  scale_colour_brewer(palette = "Dark2",
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_tufte()
-
-```
-
-The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
-ABlmed was calculated using precaculated HUD indexes.
-Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
-Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
-
-It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
-This is liekly because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
-However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
-This is not shown in the graphs, which indicate that the income levels are significantly different.
-However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
-
 Conclusion:

### 2016-08-07 code/Paper.Rmd
conclusion changes

--- 2016-08-07 code/Paper.Rmd
@@
 
-Blah blah blah blah blah
+From the collected, exploratory data, it is seen that the overall conclusion drawn is that housing affordability has decreased over the years.
+From the data, it is seen that there were many dips and peaks coincidentally near the years of 2007 and forward, which is indicative of the financial housing bubble crisis that occurred in 2007.
+
+In several of the graphs, the dips occur near 2007, 2008, and we conclude these as being consequences of the crash of 2007.
+One of the greatest factors is burden, which has increased over the years.
+If we compare
+
+Because our data only extends towards 2013, we don't have more recent data to make any great assumptions on the future of the housing market.
+According to an "ARTICLE HERE", the United States housing sector might be headed towards another bubble.

### 2016-08-07 eda/Exploration_process.Rmd
Add exploration explanation

---
title: "Exploration process"
author: "Nikolai Myllymäki"
date: "7 August 2016"
output: html_document
---
```{r, include=F}
library(ggplot2)
library(readr)
library(stringr)
library(dplyr)

clean_sample <- read_csv("../clean_data/clean_sample.csv")
clean_years <- read_csv("../clean_data/clean_years.csv")
```
Disclaimer: the plots in this document have been slightly modified to allow compilation with the modern data. Some variable names have been changed and minor settings (e.g. setting na.rm=T) have been modified. Samples of original code can be seen in /originals, but these cannot be directly compiled without modification.

As our dataset was large, with a hundred variables and nearly a million observations, the task of exploring was not trivial. We made a sample of the data to make working with it easier. Early on we decided to use a shiny app for exploration, as this would allow us to quickly change grouping and axis variables to gain insights from the dataset. Before doing the shiny app, we wanted to test the data combining and cleaning code with a simple ggplot:

```{r, echo = F}
clean_sample %>% 
  group_by(year) %>% 
  summarise(burden_year = mean(burden, na.rm = T)) %>% 
  ggplot(aes(x = year, y = burden_year)) + 
  geom_line()
```

Here we saw that using the mean was greatly skewing the results, as "sane" values for burden (housing costs divided by income) are typically between 0 and 1 but people with low income can get absurdly high values for burden as the denominator approaches zero. So we updated the plot to use medians, which are robust to outliers:

```{r , echo = F}
clean_sample %>% 
  group_by(year) %>% 
  # Take the median instead of mean to avoid extreme observations skewing the
  # result.
  summarise(burden_year = median(burden, na.rm = T)) %>% 
  ggplot(aes(x = year, y = burden_year)) + 
  geom_line()

# The housing price bubble can be seen
## But we want to analyze the effect of falling incomes as well. To be expanded.
```

Now we could see that we had at leas the rudiments of data combining and cleaning working, and we could expand on this particular graph to the shiny app. The graph looked rough because of the small sample used, but using the whole dataset smoothed things out:

```{r, echo=F}
clean_years %>% 
  group_by(year) %>% 
  # Take the median instead of mean to avoid extreme observations skewing the
  # result.
  summarise(burden_year = median(burden, na.rm = T)) %>% 
  ggplot(aes(x = year, y = burden_year)) + 
  geom_line()

# The housing price bubble can be seen
## But we want to analyze the effect of falling incomes as well. To be expanded.
```

Next we wanted to include the first version of the shiny app. This was very much a tool, since it still needed manual loading of data and didn't have options for changing the y axis variable. Still, using the app for exploring variables was very beneficial. We found out that many of the grouping variables where numeric where they should have been factors or strings, and updated the data cleaning file accordingly. We also decided on the next production plots on the ideas from the shiny-powered brainstorming session. The first version of the Shiny app can be found in the folder shiny1.

```{r, echo=F}
clean_sample %>%
  group_by(year, region) %>%  
  # Take the median instead of mean to avoid extreme observations skewing the 
  # result.
  summarise(new_y = median(aplmed, na.rm = T)) %>%
  ggplot(aes(x = year,   
                  y = new_y,   
                  colour = region)) + geom_line() +
  labs(x = "year", y = "Median Adjusted Income (per persons)", 
       title = "APLmed by region")
```

We added plots and ran into some column type issues that were addressed in the data parsing pipeline. For example in the graph above, region had to be converted from numeric to a factor. Mostly, our experimenting was done on the Shiny app, so there are not too many old graphs to begin with: most ggplot code was developed very much like it is currently on the paper. After doing different versions of time series based on the shiny app, we felt that there are aspects of the data that required different types of plots. We did the map, which was very similar to what can be seen on the paper right now, and a couple of bar charts showing burden with or without assistance.

```{r, echo = F, message=F, error=F, warning=FALSE}
clean_sample %>% filter(assisted == 0) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(y = "new_y",   
                        x = "fmtcostmedrelamicat",
                        fill = "fmtcostmedrelamicat")) +  
      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost without assistance",
       fill = "Cost as Percentage of Area Median Income",
       y="Burden",
       x = "") + 
  scale_y_continuous(limits=c(0,1)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_flip()

clean_sample %>% filter(assisted == 1) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>%  
      # Use NSE
      ggplot(aes_string(y = "new_y",   
                        x = "fmtcostmedrelamicat",
                        fill = "fmtcostmedrelamicat")) +  
      geom_bar(stat="identity") + labs(title="Burden based on Median housing cost with assistance",
       fill = "Cost as Percentage of Area Median Income",
       y="Burden",
       x = "") + 
  scale_y_continuous(limits=c(0,1)) +
  guides(fill = guide_legend(reverse = TRUE)) +
  coord_flip()
```

We received very good feedback that the colour scheme on the bar charts was a bit jarring, and that readability could be improved with a stacked version of the plot. We changed accordingly.

```{r, echo = F, message=F, error=F, warning=FALSE}
assisted = clean_sample %>% filter(!is.na(assisted)) %>% 
      # Use NSE
      group_by_("assisted", "fmtcostmedrelamicat") %>%  
      # Take the median instead of mean to avoid extreme observations skewing the 
      # result.
      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
      ungroup %>% 
      mutate(assisted = as.factor(assisted))
ggplot(assisted, aes(y = new_y,   
                  x = fmtcostmedrelamicat,
                  fill = assisted, group = assisted)) +  
  geom_bar(
           stat="identity") + 
  labs(title="Burden based on Median housing cost with assistance",
       y="Burden",
       x = "",
       fill="Assistance") + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_manual("Assistance",labels=c("Without Assistance", "With Assistance"), values=c("black", "white")) +
  coord_flip()
```

Colours were not rainbows anymore, but we still modified these by trying a bunch of different colour schemes from colorbrewer2.org. We updated the graphs in the paper to the clean tufte theme, which also happens to have the same font as default latex article output. We also updated the shiny app to include more modifiable variables, as can be seen from the final version.

### 2016-08-07 eda/originals/ablmed_region_plot.R

@@ -1,0 +1,18 @@
+library(ggplot2)
+library(readr)
+library(stringr)
+library(dplyr)
+#col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv") #col_types = col)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+             y = new_y,   
+             colour = region)) + geom_line() +
+  labs(x = "year", y = "Median Adjusted Income (per persons)", 
+       title = "ABLmed by region")

### 2016-08-07 paper/Paper.Rmd
Changed labels and removed read_csv

@@
 title: "Paper TITLE"
-output: pdf_document
+output: 
+  pdf_document: 
+    fig_caption: yes
+    latex_engine: xelatex
 ---
@@
 ```{r, echo = F}
-col_types <- "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
-
-clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+clean_years <- read_csv("../clean_data/clean_years.csv")
 
@@
     scale_colour_brewer(palette = "Dark2") +
+  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
     theme_tufte()
@@
 median_incomes <- clean_years %>%
+  mutate(region = as.character(region)) %>%
   group_by(region) %>%
@@
                                  direction = 1,
-                                 guide = "legend")
+                                 guide = "legend",
+                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})
 ```
@@
 ```{r, echo = FALSE}
-counts <- clean_years %>% 
+clean_years %>% 
   # Group and summarise the amount of people in each poverty group each year
@@
                                              "> 200% of poverty income"),
-                                  ordered = T))
-
-ggplot(counts, aes(x = year, 
+                                  ordered = T)) %>% 
+  ggplot(aes(x = year, 
              y = percent,
@@
        x = "Year",
-       y = "Precent of households",
+       y = "Percent of households",
        fill = "Household income
 relative to poverty income") +
+  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 ```{r, echo = FALSE}
-
-cleanyears <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
-  select(zinc2, year, region, fmtburden)%>% 
-  filter(fmtburden != ".") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
-
 ###### Household Income
   ####overall graph
-cleanyears %>%  
+clean_years %>% 
+  select(zinc2, year, region, fmtburden)%>% 
+  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
   group_by_("year", "fmtburden") %>%  
@@
   paste0("~ median(","zinc2", ",na.rm=T)"))) %>% 
-
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
   geom_line(size = 1) + 
-  labs(title = "Figure : Relationship between Household \nIncome and Burden", 
+  labs(title = "Relationship between Household \nIncome and Burden", 
        y = "Household Income", x = "Year", color = "Burden") +
-  scale_y_continuous(labels = unit_format(unit = "K")) +
+  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo = FALSE}
-
+```{r, echo = FALSE, fig.cap=" "}
  ### faceted by region 
-  cleanyears %>%  
+clean_years %>%  
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>%
+  filter(fmtburden != "4 No Income") %>% 
   group_by_("year","region", "fmtburden") %>%  
@@
     paste0("~ median(","zinc2", ",na.rm=T)"))) %>% 
-
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
@@
        y = "Household Income ($)", x = "Year", color = "Burden") +
+  scale_y_continuous(limits=c(0,NA),breaks=c(0,25000,50000,75000),labels=function(x){paste0(x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-cleanyearsmonthly <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
+cleanyearsmonthly <- clean_years %>% 
   select(year, zsmhc, region, fmtburden) %>% 
@@
 ####overall graph
-
 cleanyearsmonthly %>%  
-
   group_by_("year", "fmtburden") %>%  
@@
   cleanyearsmonthly %>%  
-
   group_by_("year","region", "fmtburden") %>%  
@@
     paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
-
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
@@
 ```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE}
-
-clean_sample <- read_csv("../data/clean_years.csv")
-clean_sample %>% filter(fmtstructuretype != ".") %>% 
+clean_years %>% 
+  filter(fmtstructuretype != ".") %>% 
       # Use NSE
@@
   geom_line(size = 1) +
+  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
   labs(title="Income compared to area's median grouped by house type",
@@
 ```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
-assisted = clean_sample %>% filter(!is.na(assisted)) %>% 
+assisted = clean_years %>% filter(!is.na(assisted)) %>% 
       # Use NSE
@@
        fill="Assistance") + 
-  scale_y_continuous(limits=c(0,1)) +
+  scale_y_continuous(limits=c(0,1), labels=function(x){paste0(x*100, "%")}) +
+  scale_x_discrete(labels = function(x){str_replace_all(x,"^[0-9](.*)","\\1")}) +
   scale_fill_brewer("Assistance", 
@@
 ```{r, echo = FALSE}
-
-clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
-
 clean_years %>%
@@
   summarise(new_y = median(aplmed, na.rm = T)) %>%
+  mutate(region=as.character(region)) %>% 
   ggplot(aes(x = year,   
@@
        title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
   scale_colour_brewer(palette = "Dark2", 
@@
 ```{r, echo = F}
-
-clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
-
 clean_years %>%
@@
   summarise(new_y = median(ablmed, na.rm = T)) %>%
+  mutate(region = as.character(region)) %>% 
   ggplot(aes(x = year,   
@@
        title = "Median Income Adjusted \nfor Number of Bedrooms by region") +
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
   scale_colour_brewer(palette = "Dark2",

### 2016-08-08 eda/shiny_final/ui.R
Embed shiny apps to exploration Rmd

@@ -1,6 +1,6 @@
 library(readr)
 
-clean_sample <- read_csv("../clean_data/clean_years.csv")
+clean_sample <- read_csv("../../clean_data/clean_years.csv")
 
 shinyUI(fluidPage(
 

### 2016-08-08 code/Paper_edit.Rmd
eidts paper reorganize

--- 2016-08-07 code/Paper.Rmd
@@
 ---
-title: "Paper TITLE"
+title: "Exploring Housing Affordiability in America"
 output: pdf_document
@@
 
-Summary/Intro
+##Introduction
 
@@
 
-We chose to analyze the data and ultimately view determines whether housing affordability has increased or decreased over the last few decades.
+We chose to analyze the data and ultimately determine whether housing affordability has increased or decreased over the last few decades.
 Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
-The United State’s growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
+The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
 Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
-Between all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the housing market affordability status.
+Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
-Cleaning data
+#Abstract
+
+###Cleaning data
 
@@
 Finally, the data was written to a clean_years csv file.
+
+###Analyzing the Data
+
+The main metric that is referred to in this paper is burden.
+Burden is defined by the proportion of household income that is spent towards the cost of housing.
+This metric is then further dissected by analyzing against various variables and collating data on contributing factors, such as income.
+The first data explored is income data across the four designated regions in the United States (Northeast, Midwest, South, West).
+The adjusted incomes (per persons and per bedroom) are also considered, as certain regions may have different averages for household and property sizes, which could skew the unadjusted income data.
+The behavior of household income, bracketed in ranges in accordance to census data, over time is displayed to provide a better understanding of the changes in burden over time.
+The household incomes are then compared versus burden over time.
+here it is found that the ghiher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
+Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
+finally, to elucidate the gap in burden lvels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
+
+A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
+This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
+
+###Graphical findings
+
+We start off by measuring the housing burden levels across the United States:
 
@@
 
-The graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
+This graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
 The median burden varies dramatically between households of different sizes.
@@
 Our hypothesis is that this is due to a large proportion of two-earner couples in this group.
-As housing cost burden is simply a household’s monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
+As housing cost burden is simply a household's monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
 In households of 3 or more people, the burden picks up again.
@@
 ```
+
 This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
@@
 It's interesting to see that all three burden lines are close to one another, which illustrates the principle that those with more money aren't buying more expensive houses.
-From the previous grpah, we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
+From the previous graph, we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
 Each level of burden has similar housing costs, therefore reinforcing the idea that people are buying similar priced housing among the lower and middle classes.
@@
 
-Conclusion:
+##Conclusion:
 
-Blah blah blah blah blah
+The findings in this paper help elucidate income and househodl burden trends overtime.
+However, there are many holes in the actual data.
+The data grabbed uses gross incomes and groups data over large regions of the United States.
+While the spreads of incomes and housing costs may be low, the internal varation within each region will be fairly high.
+As a result, this does not help protray a full picture of burden level across the United States, namely in (sub)urban versus non urban areas.
+Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.

### 2016-08-08 clean_data/CleanData.R

@@ -1,0 +1,147 @@
+# Running this code may take up to ten minutes because of the large size
+# of the datasets. Running from terminal instead of RStudio
+# ("Rscript CleanData.R") or on a powerful server is recommended.
+
+library(readr)
+library(dplyr)
+library(stringr)
+library(tidyr)
+
+MutateYears <- function(data, year) {
+  # Simply marks the dataset with the year to categorize the original data.
+  mutate(data, 
+         YEAR = year)
+}
+
+UpperCaseColNames <- function(frame) {
+  # Years vary over the years, so this converts the column 
+  # names to the same style as that of the latest dataset.
+  setNames(frame, 
+           (toupper(names(frame)) %>% 
+              str_replace("^COSTMED(.+)",
+                          "COSTMed\\1")))
+}
+
+AddMissingCols <- function(frame) {
+  # Datasets before 2003 are missing the three columns below, 
+  # so this function adds them to unify all of the datasets
+  mutate(frame, 
+         L30 = 0.6 * L50, 
+         GL30 = 0.6 * GL50, 
+         ABL30 = 0.6 * ABL50)
+}
+
+RenameCols <- function(frame) {
+  # Naming convention changed after the 1995 dataset, 
+  # so we match the column names.
+  rename(frame, 
+         "AGE1" = `AGE`, 
+         "FMTMETRO3" = `FMTMETRO`, 
+         "STATUS" = `ISTATUS`, 
+         "METRO3" = `METRO`)
+}
+
+base_files <- paste("../raw_data/hads",          # Loading the most recent datasets 
+                c(seq(2003, 2013, by = 2)), # first to give reference for other
+                ".bz2",                     # two sets.
+                sep = "") %>%  
+  lapply(read_csv) %>% 
+  lapply(data.frame,  # Have to convert to data frames as lapply seems 
+         stringsAsFactors = F) #to recognize csv output as named lists.
+
+base_files <- mapply(MutateYears, # Need to distinguish different years
+                 base_files,     #for datasets, so add column to do just that.
+                 as.list(seq(2003, 2013, by = 2)))
+
+modified_year <- base_files[[4]] %>% # The 2009 dataset has an extra unspecified
+  select(-VCHRMOV)                  # column that we need to remove.
+modern_files <- rbind(do.call(rbind, # Combine all of the recent years into one.
+                            base_files[c(1, 2, 3, 5, 6)]), 
[…]

### 2016-08-08 code/parse_data.R

@@ -1,33 +1,0 @@
-library(readr)
-library(dplyr)
-library(stringr)
-library(tidyr)
-
-combined_years <- read_csv("../data/combined_years.csv",
-                           # Add col_types manually to avoid parsing failures
-                           col_types = "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd",
-                           na = c("","NA", "-9", "-6", "-7", "'.'"))
-
-NoQuotations <- function(x) {
-  str_replace_all(x, "^\\'|\\'$", "")
-}
-
-# Cleaning the data
-clean_years <- combined_years %>% 
-  setNames(str_to_lower(names(.))) %>%  
-  mutate_if(is.character, NoQuotations) %>%  
-  # Clean ownrent
-  mutate(ownrent = factor(ownrent,   
-                          levels = c("1", "2"), 
-                          labels = c("own", "rent")))
-
-# Free some memory
-remove(combined_years)
-
-# Make a sample for faster developing etc.
-set.seed(1337)  
-clean_sample <- sample_n(clean_years, 1000)
-clean_years = clean_years[sample(nrow(clean_years)),]
-# Write csv files
-write_csv(clean_years, path = "../data/clean_years.csv")
-write_csv(clean_sample, path = "../data/clean_sample.csv")

### 2016-08-08 presentation/index.Rmd
Add presentation

---
title       : Housing Affordability Over the Years
subtitle    : An Exploration into America's Housing Sector
author      : Nikolai, William, Sri, Lamb
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r, echo = F, message = F}

library(ggplot2)
library(dplyr)
library(readr)
library(stringr)
library(xml2)
library(rvest)
library(tidyr)
library(purrr)
library(scales)
library(maps)
library(ggthemes)

```

```{r, echo = F}

clean_years <- read_csv("..//clean_years.csv")
regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")

```

## Introduction: Motivation and Dataset

* Our data extends from

* William's cleaning
* Sri's cleaning
* Plots
* Conclusion stuff

---

## Cleaning the Data

*

---

## Variables

Variables According to the HUD Document Data Set

Variables  | Meanings
---------- | --------
Burden | Household’s monthly housing cost divided by its monthly income
Per | Household Size 
ABLMED | Median Income Adjusted for # of Bedrooms
APLMED | Median Income Adjusted for # of Persons
ZSMHC | Monthly housing costs
ZINC2 | Household Income
Assisted | Asssisted Housing
Year | The Year of the Data
Region | Census Region
fmtincrelpovcat |HH Income Relative to Poverty Income 
fmtstructuretype | Structure Type
increlamipct | HH Income relative to AMI (Percent)
fmtcostmedrelamicat | CostMed Relative to Median Income

---

## Household Income and Burden

```{r}

clean_years %>% 
  select(zinc2, year, region, fmtburden)%>% 
  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
  group_by_("year", "fmtburden") %>%  
  summarise_(new_y = as.formula(
  paste0("~ median(","zinc2", ",na.rm=T)"))) %>%

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

```

---

## Monthly Housing Costs and Burden

```{r, echo = F, fig.align = 'center', fig.width = 12}

cleanyearsmonthly <- clean_years %>% 
  select(year, zsmhc, region, fmtburden) %>% 
  filter(fmtburden != ".") %>% 
  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))

####overall graph
cleanyearsmonthly %>%  
  group_by_("year", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte(base_size = 20)

```

---

## Regional Monthly Housing Costs and Burden

```{r}

  cleanyearsmonthly %>%  
  group_by_("year","region", "fmtburden") %>%  
  summarise_(new_y = as.formula(
    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

```

---

## Median Income Among Differing Census Regions in 2013

```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}

url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
html <- read_html(url)
xpath <- '//table[@class="tableizer-table"]'

regions_tbl <- html %>% 
  html_nodes(xpath = xpath) %>%
  html_table %>%
  .[[1]]

region_1 <- regions_tbl[2:7,] %>% 
  gather(key = region, value = state, 1:2) %>% 
  slice(1:9) %>%
  mutate(region = 1)

region_2 <- regions_tbl[10:15,] %>% 
  gather(key = region, value = state, 1:2) %>% 
  mutate(region = 2)

# Add missing Missouri
region_2[6, 2] <- "Missouri"

region_3 <- regions_tbl[18:27,] %>%  
  gather(key = region, value = state, 1:2) %>% 
  slice(c(-10, -15, -16)) %>%
  mutate(region = 3)

region_4 <- 
  regions_tbl[30:37,] %>%  
  gather(key = region, value = state, 1:2) %>% 
  slice(1:13) %>%
  mutate(region = 4)

regions <- rbind(region_1, region_2, region_3, region_4) %>%
  mutate(state = str_to_title(state),
         region = as.character(region))

median_incomes <- clean_years %>%
  filter(year = 2013) %>%
  mutate(region = as.character(region)) %>%
  group_by(region) %>%
  summarise(zinc2_median = median(zinc2, na.rm = T))

map_df <- map_data("state") %>%
  mutate(region = str_to_title(region)) %>%
  rename(state = region)  %>%
  # Join with census region information
  left_join(regions) %>%
  left_join(median_incomes)

map_df %>% 
  ggplot() +  
  geom_polygon(aes(x=long, 
                   y = lat, 
                   fill = zinc2_median, 
                   group = group)) + 
  coord_fixed(1.3) +
  theme_tufte() +
  scale_x_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  scale_y_continuous(name = NULL,
                     labels = NULL,
                     breaks = NULL) +
  labs(fill = "Median income") +
  scale_fill_distiller(limits = c(25000, 42000), 
                                 breaks = sort(unique(map_df$zinc2_median),
                                               decreasing = T),
                                 palette = "Greens", 
                                 direction = 1,
                                 guide = "legend",
                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})

```

---

## Household Size and Burden

```{r, echo = F}

  clean_years %>%
    mutate(per = cut(per, 
                     breaks = c(0, 1, 2, Inf),
                     labels = c("1", "2", "3 or more")))  %>%
    group_by(year, per) %>%  
    # Take the median instead of mean to avoid extreme observations skewing the 
    # result.
    summarise(new_y = median(burden, na.rm = T)) %>%
    ggplot(aes(x = year,   
                      y = new_y,   
                      colour = per)) +  
    geom_line(size = 1) +
    labs(colour = "Household Size",
         x = "Year",
         y = "Median Burden",
         title = "Relationship between Burden and Household Size") +
    scale_colour_brewer(palette = "Dark2") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
    theme_tufte()

```

---

## Households and Poverty

```{r, echo = F, fig.align = 'center', fig.width = 12}
clean_years %>% 
  # Group and summarise the amount of people in each poverty group each year
  group_by(fmtincrelpovcat, year) %>%
  summarise(count = n()) %>%
  # Clean NAs
  na.omit() %>%
  # Group again by year to calculate relative percentages per year
  group_by(year) %>% 
  mutate(percent = count / sum(count),  
         # Clean the labels
         fmtincrelpovcat = factor(fmtincrelpovcat, 
                                  levels = c("1 LTE Poverty", 
                                             "2 100-150% Poverty", 
                                             "3 150-200% Poverty", 
                                             "4 200%+ Poverty"),  
                                  labels = c("Less than poverty income",   
                                             "100-150% of poverty income",   
                                             "150-200% of poverty income",    
                                             "> 200% of poverty income"),
                                  ordered = T)) %>% 
  ggplot(aes(x = year, 
             y = percent,
             fill = fmtincrelpovcat)) + 
  geom_area(alpha = 0.8) +
  # Use a nice colour palette
  scale_fill_brewer(palette = "RdYlGn", 
                    # Reverse the order of legend items to match the
                    # stacking order
                    guide = guide_legend(reverse=TRUE)) +
  labs(title = "Number of Households Relative to the Poverty Line",
       x = "Year",
       y = "Percent of Households",
       fill = "Household Income \nRelative to Poverty Income") +
  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()
```

---

## MORE GRAPHS

```{r, echo = F, fig.align = 'center', fig.width = 12}

clean_years %>% 
  select(zinc2, year, region, fmtburden)%>% 
  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
  group_by_("year", "fmtburden") %>%  
  summarise_(new_y = as.formula(
  paste0("~ median(","zinc2", ",na.rm=T)"))) %>%

  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
  geom_line(size = 1) + 
  labs(title = "Relationship between Household \nIncome and Burden", 
       y = "Household Income", x = "Year", color = "Burden") +
  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
  scale_colour_brewer(palette = "Dark2") +
  theme_tufte()

```

---

## Concluding Thoughts

* Housing Affordability Has Decreased

* The 2007 Housing Bubble Crisis

* Different Housing Regions Affected Differently

  * Western and Northeast Regions Vs. South and Midwest

---

## Future Questions

* Does the Data in some graphs keep showing a downward slope?

* If we analyzed data from 2013 - 2016, would it show another potential, housing housing bubble?

* Our data is limited to 2013, so what is the housing situation most recently?

---

## Ask Us Anything

---

### 2016-08-08 presentation/index.Rmd
presentation changes

@@
 
-clean_years <- read_csv("..//clean_years.csv")
+clean_years <- read_csv("../clean_data/clean_years.csv")
 regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
@@
 median_incomes <- clean_years %>%
-  filter(year = 2013) %>%
+  ##filter(year == 2013) %>%
   mutate(region = as.character(region)) %>%

### 2016-08-08 code/Paper_edit.Rmd
edits
minor edits

@@
 ---
-title: "Exploring Housing Affordiability in America"
+title: "Exploring United States Housing Affordability"
 output: pdf_document
@@
 Then, the combined years dataset was cleaned using the piping methods in dplyr.
-The data was rather messy, with unecessary quotaions around character values and inconsistent letter cases in the column names.
+The data was rather messy, with unecessary quotations around character values and inconsistent letter cases in the column names.
 This was rectified by setting all the column names to lowercase and removing all the quotations from character values.
@@
 
-It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
-This is liekly because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+It is interesting to note that the disparity between the incomes increases between the Northeast and Western regions and the Midwest and South.
+This is likely because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
 However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
 This is not shown in the graphs, which indicate that the income levels are significantly different.
-However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
+However, if housing affordability was factored into the calculation, the income disparity would be far less if not almost nonexistent.
 

### 2016-08-08 presentation/index.Rmd
Fix map

@@
 median_incomes <- clean_years %>%
-  ##filter(year == 2013) %>%
+  filter(year == 2013) %>%
   mutate(region = as.character(region)) %>%
@@
 
-map_df <- map_data("state") %>%
+us_map <- map_data("state") %>%
   mutate(region = str_to_title(region)) %>%
@@
 
-map_df %>% 
+us_map %>% 
   ggplot() +  
@@
   labs(fill = "Median income") +
-  scale_fill_distiller(limits = c(25000, 42000), 
-                                 breaks = sort(unique(map_df$zinc2_median),
+  scale_fill_distiller(limits = c(35000, 54000), 
+                                 breaks = sort(unique(us_map$zinc2_median),
                                                decreasing = T),
@@
                         labels=function(x){paste0("$",x %/% 100 / 10, "K")})
-
 ```

### 2016-08-08 presentation/index.Rmd
fixes

--- 2016-08-08 presentation/index.Rmd
@@
 
-*
+Combined Years:
+
+- Dataset is split into fifteen datasets, one for each odd year between 1985 and 2013
+- To standardize the data used data from 2003 - 2013
+- Reorder the columns to exactly match that of our base (2003 onward)
+- Combine every list of datasets into single dataset
+
+Clean Data:
+
+- Initally removed NAs, expressed column in read_csv
+- Set column names to lowercase, removed quotations
+- Created column for owning and renting data (1= own, 2= rent)
+- Write to csv (same for combined years)
 
@@
 
-## Variables
-
-Variables According to the HUD Document Data Set
+## Variables According to the HUD Document Data Set
 
@@
 
-```{r}
+```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}
 
@@
   filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
+  mutate(fmtburden = str_replace(fmtburden, 
+                                 "Greater than 50%", 
+                                 "50% or More")) %>% 
   group_by_("year", "fmtburden") %>%  
@@
        y = "Household Income", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
+  scale_y_continuous(limits=c(0,NA),
+                     breaks=c(0,20000,40000,60000),
+                     labels=function(x){paste0(x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  theme_tufte(base_size = 20)
 
@@
 
-```{r}
+```{r, echo = F, fig.align = 'center', fig.width = 12, message = F }
 
@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  theme_tufte(base_size = 20)
 
@@
   coord_fixed(1.3) +
-  theme_tufte() +
+  theme_tufte(base_size = 20) +
   scale_x_continuous(name = NULL,
@@
   scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
-    theme_tufte()
+    theme_tufte(base_size = 20)
 
@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  theme_tufte(base_size = 20)
+```
+
+---
+
+## Income Compared to Area's Median by House TypeS
+
+```{r, echo = F, fig.align = 'center', fig.width = 12, message =F}
+
+clean_years %>% filter(fmtstructuretype != ".") %>% 
+      # Use NSE
+      group_by_("year", "fmtstructuretype") %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
+      # Use NSE
+      ggplot(aes_string(x = "year",   
+                        y = "new_y",   
+                        colour = "fmtstructuretype")) +  
+  geom_line(size = 1) + 
+  labs(colour = "Structure Type",
+       y="Income percentage of area median income",
+       x = "Year") +
+  scale_colour_brewer(palette = "Dark2") +
+  theme_tufte(base_size = 20)
+
+```
+
+---
+
+## Burden Based on Median Housing Cost with Assistance
+
+```{r, echo = F, fig.align = 'center', fig.width = 12, message= F, warning=F}
+
+assisted = clean_years %>% filter(!is.na(assisted)) %>% 
+      # Use NSE
+      group_by_("assisted", "fmtcostmedrelamicat") %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
+      ungroup %>% 
+      mutate(assisted = as.factor(assisted))
+ggplot(assisted, aes(y = new_y,   
+                  x = fmtcostmedrelamicat,
+                  fill = assisted, group = assisted)) +  
+  geom_bar(
+           stat="identity") + 
+  labs(y="Burden",
+       x = "",
+       fill="Assistance") + 
+  scale_y_continuous(limits=c(0,1)) +
+  scale_fill_brewer("Assistance", 
+                    labels = c("Without Assistance", 
+                               "With Assistance"),
+                    palette = "Set2")  +
+  coord_flip() +
+  theme_tufte(base_size = 20)
+
 ```
@@
 
-```{r, echo = F, fig.align = 'center', fig.width = 12}
+```{r, echo = F, fig.align = 'center', fig.width = 12, message= F}
 
 clean_years %>%
-  select(zinc2, year, region, fmtburden)%>% 
-  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
-  group_by_("year", "fmtburden") %>%  
-  summarise_(new_y = as.formula(
-  paste0("~ median(","zinc2", ",na.rm=T)"))) %>%
+  mutate(region = as.character(region)) %>% 
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
+       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte(base_size = 20)
+```
 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+- Adjusted income per persons
+- Bracketed by HUD indexes (0.70 for 1 person, 0.80 for 2, 1.00 for 4)
+- Spike around 2010 associated with financial crisis
+
+---
+
+## Median Income Adjusted for Number of Bedrooms By Region
+
+```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}
+
+clean_years %>%
+  mutate(region = as.character(region)) %>% 
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
   geom_line(size = 1) +
-  labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
-  scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)") + 
+  scale_colour_brewer(palette = "Dark2",
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte(base_size = 20)
 
 ```
+
+Adjusted income per bedroom
+
+Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)
+
+More discrepancy in burden levels between the Midwest/South and Northeast/West
+
+  - Larger homes in the Midwest/South
+
+Purhasing parity shoudl be higher in midwest/south, not expressed in data
 
@@
 
-* Housing Affordability Has Decreased
+Housing Affordability Has Decreased
 
-* The 2007 Housing Bubble Crisis
+The 2007 Housing Bubble Crisis and it's Consequences
 
-* Different Housing Regions Affected Differently
+Census Regions Differences
 
-  * Western and Northeast Regions Vs. South and Midwest
+  - Western and Northeast Regions Vs. South and Midwest
 

### 2016-08-08 clean_data/CleanData.R
Remove numbers from burden and structure

@@ -124,7 +124,9 @@ clean_years <- combined_years %>%
   # Clean ownrent
   mutate(ownrent = factor(ownrent,   
                           levels = c("1", "2"), 
-                           labels = c("own", "rent"))) %>%
+                          labels = c("own", "rent")),
+         fmtburden = str_replace_all(fmtburden, "^[0-9] ", ""),
+         fmtstructuretype = str_replace_all(fmtstructuretype, "^[0-9] ", "")) %>%
   na_if("") %>%
   na_if("NA") %>%
   na_if("-5") %>%

### 2016-08-08 presentation/index.Rmd
Change plot fonts to sans serif

@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 
@@
   coord_fixed(1.3) +
-  theme_tufte(base_size = 20) +
+  theme_tufte(base_size = 20, base_family = "sans serif") +
   scale_x_continuous(name = NULL,
@@
   scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
-    theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 
@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 ```
@@
   coord_flip() +
-  theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 
@@
                       labels = regionNames) +
-  theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 ```
@@
                       labels = regionNames) +
-  theme_tufte(base_size = 20)
+  theme_tufte(base_size = 20, base_family = "sans serif")
 

### 2016-08-08 presentation/index.Rmd
slight changes

--- 2016-08-08 presentation/index.Rmd
@@
   geom_line(size = 1) + 
-  labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income", x = "Year", color = "Burden") +
+  labs(y = "Household Income", x = "Year", color = "Burden") +
   scale_y_continuous(limits=c(0,NA),
@@
   geom_line(size = 1) + 
-  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
-       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
   scale_colour_brewer(palette = "Dark2") +
@@
 ```
+
+- People using savings, not paying for costs, paying with their trust funds
 
@@
   facet_wrap(~region, labeller = as_labeller(regionNames)) + 
-  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
-       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
   scale_colour_brewer(palette = "Dark2") +
@@
 ```
+
+- The Northeast was impacted first, and then the other regions lagged by two years
 
@@
          x = "Year",
-         y = "Median Burden",
-         title = "Relationship between Burden and Household Size") +
+         y = "Median Burden") +
     scale_colour_brewer(palette = "Dark2") +
@@
 
-## Households and Poverty
+## Households Relative to the Poverty Line
 
@@
                     guide = guide_legend(reverse=TRUE)) +
-  labs(title = "Number of Households Relative to the Poverty Line",
-       x = "Year",
+  labs(x = "Year",
        y = "Percent of Households",
@@
 
-## MORE GRAPHS
+## "Median Income Adjusted for the Number of People by Region
 
@@
   geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
-       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
   scale_colour_brewer(palette = "Dark2", 
@@
                       labels = regionNames) +
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K")}) +
   theme_tufte(base_size = 20)
@@
 - Adjusted income per persons
+
 - Bracketed by HUD indexes (0.70 for 1 person, 0.80 for 2, 1.00 for 4)
+
 - Spike around 2010 associated with financial crisis
@@
 
+--- &twocol
+
 ## Median Income Adjusted for Number of Bedrooms By Region
 
-```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}
+*** =left
+
+```{r, echo = F, message = F}
 
@@
   labs(x = "Year", y = "Median Adjusted Income (Per Bedroom)") + 
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2",
@@
 
-Adjusted income per bedroom
+*** =right
 
-Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)
+- Adjusted income per bedroom
 
-More discrepancy in burden levels between the Midwest/South and Northeast/West
+- Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)
+
+- More discrepancy in burden levels between the Midwest/South and Northeast/West
 
@@
 
-Purhasing parity shoudl be higher in midwest/south, not expressed in data
+- Purhasing parity should be higher in midwest/south (not expressed in data)
 

### 2016-08-08 presentation/index.Rmd

--- 2016-08-08 presentation/index.Rmd
@@
                       labels = regionNames) +
+
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K")}) +
+
   theme_tufte(base_size = 20, base_family = "sans serif")
+
 ```

### 2016-08-08 paper/Paper.Rmd
Changed LTE in histogram plot

@@
   scale_y_continuous(limits=c(0,1), labels=function(x){paste0(x*100, "%")}) +
-  scale_x_discrete(labels = function(x){str_replace_all(x,"^[0-9](.*)","\\1")}) +
+  scale_x_discrete(labels = function(x){str_replace_all(str_replace_all(x, "^[0-9] (.*)","\\1"), "LTE (.*)","<= \\1")}) +
   scale_fill_brewer("Assistance", 

### 2016-08-08 presentation/index.Rmd
Added parsing stuff

@@
   scale_y_continuous(limits=c(0,1)) +
+  scale_x_discrete(labels = function(x){str_replace_all(str_replace_all(x, "^[0-9] (.*)","\\1"), "LTE (.*)","<= \\1")}) +
   scale_fill_brewer("Assistance", 

### 2016-08-08 presentation/index.Rmd
Final presentation version, modify .gitignore

--- 2016-08-08 presentation/index.Rmd
@@
 
-* Our data extends from
+- Data is chosen from the American Housing Survey (1985 - 2013)
 
-* William's cleaning
-* Sri's cleaning
-* Plots
-* Conclusion stuff
+- Burden = Housing Cost Over Monthly Income
+
+  - Households with zero/negative income have "-1" in data
+
+- Goal: Housing Affordability
+
+- Hypothesis: Affordability decreased due to key significant events
+
+- US growing income inequality is indicator of:
+  - Housing affordability
+  - Shifting income brackets poor and rich
+
+- 2007 Housing Market Crash
 
@@
 fmtcostmedrelamicat | CostMed Relative to Median Income 
+AMI | Area median income
 
@@
 
-```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}
+```{r, echo = F, fig.align = 'center', fig.width = 12, message = F, warning=F}
 
@@
   select(zinc2, year, region, fmtburden)%>% 
-  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
+  filter(fmtburden != ".", fmtburden != "No Income") %>% 
   mutate(fmtburden = str_replace(fmtburden, 
@@
                      breaks=c(0,20000,40000,60000),
-                     labels=function(x){paste0(x/1000, "K")}) +
+                     labels=function(x){paste0("$", x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo = F, fig.align = 'center', fig.width = 12}
+```{r, echo = F, fig.align = 'center', fig.width = 12, fig.height=6, warning = F}
 
@@
   filter(fmtburden != ".") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
-
-####overall graph
+  mutate(fmtburden = str_replace(fmtburden, 
+                                 "Greater than 50%", 
+                                 "50% or More"))
 cleanyearsmonthly %>%  
@@
 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  ggplot(aes_string(x = "year", 
+                    y = "new_y", 
+                    color = "fmtburden")) + 
   geom_line(size = 1) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(y = "Monthly Housing Costs", 
+       x = "Year", 
+       color = "Burden") +
+  scale_y_continuous(labels = function(x){paste0("$", x)}) +
+  scale_colour_brewer(palette = "Dark2") +
+  theme_tufte(base_size = 20, base_family = "sans serif")
+
+```
+- No Income Bracket
+  - Not paying for costs
+  - Paying with trust funds
+- Similar housing costs among burden brackets
+
+---
+
+## Regional Monthly Housing Costs and Burden
+
+- The Northeast impacted first
+  - Other regions lagged by two years
+```{r, echo = F, fig.align = 'center', fig.width = 12, fig.height = 6, message = F }
+
+  cleanyearsmonthly %>%  
+  group_by_("year","region", "fmtburden") %>%  
+  summarise_(new_y = as.formula(
+    paste0("~ median(","zsmhc", 
+           ",na.rm=T)"))) %>% 
+  ggplot(aes_string(x = "year", 
+                    y = "new_y", 
+                    color = "fmtburden")) + 
+  geom_line(size = 1) + 
+  facet_wrap(~region, 
+             labeller = as_labeller(regionNames)) + 
+  labs(y = "Monthly Housing Costs", 
+       x = "Year", 
+       color = "Burden") +
+  scale_y_continuous(labels = function(x){paste0("$", x)}) + 
   scale_colour_brewer(palette = "Dark2") +
@@
 
-- People using savings, not paying for costs, paying with their trust funds
-
 ---
 
-## Regional Monthly Housing Costs and Burden
+## Median Income by Census Regions in 2013
 
-```{r, echo = F, fig.align = 'center', fig.width = 12, message = F }
-
-  cleanyearsmonthly %>%  
-  group_by_("year","region", "fmtburden") %>%  
-  summarise_(new_y = as.formula(
-    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line(size = 1) + 
-  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
-  scale_colour_brewer(palette = "Dark2") +
-  theme_tufte(base_size = 20, base_family = "sans serif")
-
-```
-
-- The Northeast was impacted first, and then the other regions lagged by two years
-
----
-
-## Median Income Among Differing Census Regions in 2013
-
-```{r, echo = F, fig.align = 'center', fig.width = 12, message = F}
+```{r, echo = F, fig.align = 'center', fig.width = 14, message = F}
 
@@
 
-```{r, echo = F}
+```{r, echo = F, fig.align = 'center', fig.width = 12}
 
@@
 
-## Income Compared to Area's Median by House TypeS
+## Income of Median Area by House Types
 
@@
   labs(colour = "Structure Type",
-       y="Income percentage of area median income",
+       y="Income Percentage of Area Median Income",
        x = "Year") +
+  scale_y_continuous(labels = function(x){paste0(x,"%")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-## Burden Based on Median Housing Cost with Assistance
-
-```{r, echo = F, fig.align = 'center', fig.width = 12, message= F, warning=F}
+## Burden and Median Housing Cost with Assistance
+```{r,echo = F,fig.align='center',fig.width = 12, fig.height=4.5,message= F,warning=F}
 
@@
   theme_tufte(base_size = 20, base_family = "sans serif")
-
 ```
+- Those needing assitance with reducment of total salary
+  - Unemployment 
+  - Monetary Crisis 
+- High cost brackets = Higher burdens with assistance 
+- Less mobility for single families
 
@@
 
-## "Median Income Adjusted for the Number of People by Region
+## Median Income Adjusted by Region
 
-```{r, echo = F, fig.align = 'center', fig.width = 12, message= F}
+```{r, echo = F, fig.align = 'center', fig.width = 12, fig.height = 6 , message= F}
 
@@
   geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
+  labs(x = "Year", y = "Median Adjusted Income (Per Person)") + 
   scale_colour_brewer(palette = "Dark2", 
@@
 
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K")}) +
+  scale_y_continuous(labels=function(x){paste0("$", x/1000, "K")}) +
 
@@
 - Adjusted income per persons
-
-- Bracketed by HUD indexes (0.70 for 1 person, 0.80 for 2, 1.00 for 4)
-
-- Spike around 2010 associated with financial crisis
-
----
+- Bracketed by HUD indexes (0.70 = 1 person, 0.80 = 2, 1.00 = 4)
+- Spike around 2010
 
@@
 
-## Median Income Adjusted for Number of Bedrooms By Region
+## Median Income Adjusted for Bedrooms By Region
 
@@
 
-```{r, echo = F, message = F}
+```{r, echo = F, message = F, fig.width = 7}
 
@@
 
-- Adjusted income per bedroom
+- Bracketedby HUD indexes
 
-- Bracketedby HUD indexes (0.70 for 0, 0.90 for 2, 1.04 for 3)
+  - (0.70 for 0, 0.90 for 2, 1.04 for 3)
 
-- More discrepancy in burden levels between the Midwest/South and Northeast/West
+- More discrepancy in burden levels between regions
 
-  - Larger homes in the Midwest/South
+  - Larger homes in Midwest/South
 
-- Purhasing parity should be higher in midwest/south (not expressed in data)
+- Purchasing parity higher in midwest/south
 
@@
 
-Housing Affordability Has Decreased
+- Housing Affordability Has Decreased
 
-The 2007 Housing Bubble Crisis and it's Consequences
+- The 2007 Housing Bubble Crisis and it's Consequences
 
-Census Regions Differences
+- Census Regions Differences
 
@@
 
----
+## Future Questions / Ask Us Anything
 
-## Future Questions
+- Do these graphs continue to have decreasing slopes 2014 - Present?
 
-* Does the Data in some graphs keep showing a downward slope?
-
-* If we analyzed data from 2013 - 2016, would it show another potential, housing housing bubble?
-
-* Our data is limited to 2013, so what is the housing situation most recently?
+- If we analyzed data from 2013 - 2016, would it show another potential, housing bubble?
 
@@
 ## Ask Us Anything
-
----

### 2016-08-08 paper/Paper.Rmd
Removed titles and added figures

--- 2016-08-07 paper/Paper.Rmd
@@
 
-```{r, echo = F}
+```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
 clean_years <- read_csv("../clean_data/clean_years.csv")
@@
 
-```{r, include = F}
+```{r, include = F, fig.cap="Median Income across Census Regions"}
 url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap="Amount of households relative to poverty line"}
 clean_years %>% 
@@
                     guide = guide_legend(reverse=TRUE)) +
-  labs(title = "Amount of households relative to poverty line",
-       x = "Year",
+  labs(x = "Year",
        y = "Percent of households",
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
 ###### Household Income
@@
   geom_line(size = 1) + 
-  labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income", x = "Year", color = "Burden") +
+  labs(y = "Household Income", x = "Year", color = "Burden") +
   scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
@@
 
-```{r, echo = FALSE, fig.cap=" "}
+```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
  ### faceted by region 
@@
   facet_wrap(~region, labeller = as_labeller(regionNames)) +
-  labs(title = "Relationship between Household \nIncome and Burden", 
-       y = "Household Income ($)", x = "Year", color = "Burden") +
+  labs(y = "Household Income ($)", x = "Year", color = "Burden") +
   scale_y_continuous(limits=c(0,NA),breaks=c(0,25000,50000,75000),labels=function(x){paste0(x/1000, "K")}) +
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap = "Relationship between Monthly Housing Costs and Burden"}
 
@@
 
-```{r, echo= FALSE}
+```{r, echo= FALSE, fig.cap="Relationship between Monthly Housing Costs and Burden"}
 
@@
 
-```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE}
+```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE, fig.cap = "Income compared to area's median grouped by house type"}
 clean_years %>% 
@@
   scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
-  labs(title="Income compared to area's median grouped by house type",
-       colour = "Structure Type",
+  labs(colour = "Structure Type",
        y="Income percentage of area median income",
@@
 
-```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
+```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median housing cost with assistance"}
 assisted = clean_years %>% filter(!is.na(assisted)) %>% 
@@
            stat="identity") + 
-  labs(title="Burden based on Median housing cost with assistance",
-       y="Burden",
+  labs(y="Burden",
        x = "",
@@
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region"}
 clean_years %>%
@@
   geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
-       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
   scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
@@
 
-```{r, echo = F}
+```{r, echo = F, fig.cap="Median Income Adjusted for Number of Bedrooms by region"}
 clean_years %>%
@@
   geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
-       title = "Median Income Adjusted \nfor Number of Bedrooms by region") +
+  labs(x = "Year", y = "Median Adjusted Income (per persons)") +
   scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +

### 2016-08-08 eda/Exploration_process.Rmd

@@
 title: "Exploration process"
-author: "Nikolai Myllymäki"
-date: "7 August 2016"
 output: html_document
+runtime: shiny
 ---
@@
 
-Next we wanted to include the first version of the shiny app.
-This was very much a tool, since it still needed manual loading of data and didn't have options for changing the y axis variable.
+Next we wanted to include the first version of the shiny app. This was very much a tool, since it still needed manual loading of data and didn't have options for changing the y axis variable.
+
+```{r, echo=FALSE}
+ sidebarLayout(
+    sidebarPanel(
+
+      selectInput("sel_var",
+                  label = "Grouping variable",
+                  choices = names(clean_sample),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
+                  selected = "ownrent")
+
+      ),
+
+    mainPanel(
+    )
+  )
+
+renderPlot({
+
+     clean_sample %>%  
+      # Use NSE
+      group_by_("year", input$sel_var) %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise(burden_year = median(burden, na.rm = T)) %>%  
+      # Use NSE
+      ggplot(aes_string(x = "year",   
+                        y = "burden_year",   
+                        colour = input$sel_var)) +  
+      geom_line()
+
+  })
+```
+
 Still, using the app for exploring variables was very beneficial.
@@
 We also decided on the next production plots on the ideas from the shiny-powered brainstorming session.
-The first version of the Shiny app can be found in the folder shiny1.
+The first version of the Shiny app can also be found in the folder /shiny_first.
 
@@
 
-We added plots and ran into some column type issues that were addressed in the data parsing pipeline.
-For example in the graph above, region had to be converted from numeric to a factor.
-Mostly, our experimenting was done on the Shiny app, so there are not too many old graphs to begin with: most ggplot code was developed very much like it is currently on the paper.
+We added plots and ran into some column type issues that were addressed in the data parsing pipeline. For example in the graph above, region had to be converted from numeric to a factor.
+
+We updated the shiny app to include more modifiable variables and to use the whole dataset, as can be seen from the final version of the app.
+Mostly, our experimenting was done on the Shiny app, so there are not too many old graphs to begin with.
+Most ggplot code was developed very much like it is currently on the paper with the help of the shiny app.
+
+```{r, echo=F}
+  sidebarLayout(
+    sidebarPanel(
+
+      selectInput("sel_color",
+                  label = "Colour",
+                  choices = names(clean_years),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
+                  selected = "ownrent"),
+      selectInput("sel_y",
+                  label = "Y axis",
+                  choices = names(clean_years),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
+                  selected = "burden"),
+      selectInput("sel_x",
+                  label = "X axis",
+                  choices = names(clean_years),
+                  # Select "ownrent" as initial value to avoid getting a 
+                  # non-categorical variable by default
+                  selected = "year")
+      ),
+
+    mainPanel(
+    )
+  )
+
+renderPlot({ 
+     clean_years %>%  
+      # Use NSE
+      group_by_(input$sel_x, input$sel_color) %>%  
+      # Take the median instead of mean to avoid extreme observations skewing the 
+      # result.
+      summarise_(new_y= as.formula(paste0("~ median(",input$sel_y, ",na.rm=T)"))) %>%  
+      # Use NSE
+      ggplot(aes_string(x = input$sel_x,   
+                        y = "new_y",   
+                        colour = input$sel_color)) +  
+      geom_line()
+
+  })
+```
+
 After doing different versions of time series based on the shiny app, we felt that there are aspects of the data that required different types of plots.
@@
 
-We received very good feedback that the colour scheme on the bar charts was a bit jarring, and that readability could be improved with a stacked version of the plot. We changed accordingly.
+We received feedback that the colour scheme on the bar charts was a bit jarring, and that readability could be improved with a stacked version of the plot. We changed accordingly.
 
@@
 We updated the graphs in the paper to the clean tufte theme, which also happens to have the same font as default latex article output.
-We also updated the shiny app to include more modifiable variables, as can be seen from the final version.

### 2016-08-08 paper/Paper.Rmd
Add lines to plots

@@
   scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
-    theme_tufte()
+  theme_minimal(base_family = "serif")
 
@@
   coord_fixed(1.3) +
-  theme_tufte() +
+  theme_minimal(base_family = "serif") +
   scale_x_continuous(name = NULL,
@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  theme_minimal(base_family = "serif") 
 ```
@@
   scale_colour_brewer(palette = "Dark2") +
-  theme_tufte()
+  theme_minimal(base_family = "serif")
 
@@
   coord_flip() +
-  theme_tufte()
+  theme_minimal(base_family = "serif")
 ```
@@
                       labels = regionNames) +
-  theme_tufte()
+  theme_minimal(base_family = "serif")
 

### 2016-08-08 paper/Paper.Rmd
Sri's graphs side-by-side

@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region"}
+```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=5, fig.height=5}
 clean_years %>%
@@
   mutate(region=as.character(region)) %>% 
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_minimal(base_family = "serif")
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  mutate(region = as.character(region)) %>% 
   ggplot(aes(x = year,   

### 2016-08-08 code/Paper_edit.Rmd
rmd

@@
 
-We chose the dataset from the American Housing Survey, which derives sets of data from 1985 and on.
+  The dataset was taken from the American Housing Survey, which derives sets of data from 1985 and on.
 The data set includes categorizing each housing unit by factors such as affordability, income through Fair Market Rent and the Adjusted Median Income.
@@
 Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
-
-We chose to analyze the data and ultimately determine whether housing affordability has increased or decreased over the last few decades.
+  The goal is to determine whether housing affordability has increased or decreased over the last few decades.
 Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
@@
 Finally we combine every list of datasets into a single dataset, and then store it in a comma separated file called combined_years.
-
   The combined_years data was read into a dataframe using the read_csv function.
 Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
-Also, the nonsensical integer date (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
+Also, the nonsensical integers (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
 Then, the combined years dataset was cleaned using the piping methods in dplyr.
@@
 Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
-finally, to elucidate the gap in burden lvels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
 Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
-
   A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
+In particular, the lag in the downward spike after 2007 between the Northeast and the other regions is discussed.
 This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
@@
 This is understandable, since people living alone cannot split any costs between other earners.
-
   Surprisingly, households of two people have the lowest burdens overall.
@@
 This is likely due to a larger amount of children and / or retirees in larger households.
-
+Also, households with 3 or more members tend to hae only one earner.
   Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007.
@@
 We used median incomes instead of mean incomes to avoid skewing the results with outliers.
-
   The household median monthly incomes vary between different census regions from 32200 dollars in the South to nearly 25% higher, 40000, in the West.
 The Northeast trails closely behind the West with 37267.50 dollars, and the Midwest is closer to the South with a median income of 35000.
-
   While the dataset doesn't have more granular location data, even these four regions are very distinct from each other.
@@
 
-This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+  In order to better understand the regional differences in income, the incomes were adjusted by certain variables. 
+  The first plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
 APLmed is the adjusted median household income per person.
@@
 The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
-
-It is clear that the Northeast and the West are the most expensive regions. This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
-
+  It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
   The median adjusted income increases steadily (roughly linearly) over time.
@@
 Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
-
   It is interesting to note that the disparity between the incomes increases between the Northeast and Western regions and the Midwest and South.
@@
 However, if housing affordability was factored into the calculation, the income disparity would be far less if not almost nonexistent.
-
 The lag in the downward trend after 2008 still occurs in the plots.
+  With the income information in hand, the burden data can be explored in further detail.
 
@@
 Each household has been calculated a poverty treshold according to the Census Bureau’s official thresholds, based on family size and number of children.
-
   As can be seen from the plot, the amount of people in different income groups stayed relatively constant between 1985 and 2007.
@@
 This stable relationship broke noticeably in the financial crisis of 2007 and the following recession, when the amount of people in poverty started a quite steep and steady climb upwards.
-
   Interestingly, the income brackets of 100%-200% stayed almost exactly the same size as before, while the increase in poverty was in effect solely the result of the dwindling of the highest (over 200% of poverty income) income bracket.
@@
 
-The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
+  Further delving into the behavior of burden amongst the lower income workers, this plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
 At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance.
@@
 Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
+  Further analysis could involve predicting the trends for the burden and income data to corroborate whether or not a two year lag exists between the Northeast and the other regions following the financial meltdown.
+Also, adjusting income by cost of living would provide a better understanding of burden amongst differing regions.
+Regions could be further subsetted by town or district level to provide more granular data.

### 2016-08-08 paper/Paper.Rmd
rmd

--- 2016-08-08 paper/Paper.Rmd
@@
 ---
-title: "Paper TITLE"
-output: 
-  pdf_document: 
-    fig_caption: yes
-    latex_engine: xelatex
+title: "Exploring Housing Affordiability in America"
+output: pdf_document
 ---
@@
 Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+  We chose to analyze the data and ultimately determine whether housing affordability has increased or decreased over the last few decades.
+Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
+The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
+Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
+Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
-We chose to analyze the data and ultimately view determines whether housing affordability has increased or decreased over the last few decades.
-Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
-The United State’s growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
-Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
-Between all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the housing market affordability status.
+#Abstract
 
-Cleaning data
+###Cleaning data
 
-Cleaning data
   Initially, the dataset is split into fifteen datasets, one for each odd year between 1985 and 2013.
@@
 Finally we combine every list of datasets into a single dataset, and then store it in a comma separated file called combined_years.
+  The combined_years data was read into a dataframe using the read_csv function.
+Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
+Also, the nonsensical integer date (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
+Then, the combined years dataset was cleaned using the piping methods in dplyr.
+The data was rather messy, with unecessary quotaions around character values and inconsistent letter cases in the column names.
+This was rectified by setting all the column names to lowercase and removing all the quotations from character values.
+One the main premises of this paper is comparing the burden of household that are owning and renting.
+Since there was no explicit labeling of data corresponding to owning and renting households, an  additional column "ownrent" was created which labeled levels 1 and 2 to own and rent as factors.
+Finally, the data was written to a clean_years csv file.
 
-The "combined_ years" data was read into a dataframe using the "read_ csv" function.
-Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), and the nonsensical integer date (negative numbers) were converted to NAs in the dataset.
-Then, the combined years dataset was cleaned using the piping methods in dplyr.
-All the column names were set to lowercase, and all the quotations were removed from character values.
-Then, an additional column “ownrent” was added which labeled levels 1 and 2 to own and rent as factors. Then, the data was written to a clean_years csv file.
+###Analyzing the Data
 
-```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
-clean_years <- read_csv("../clean_data/clean_years.csv")
+  The main metric that is referred to in this paper is burden.
+Burden is defined by the proportion of household income that is spent towards the cost of housing.
+This metric is then further dissected by analyzing against various variables and collating data on contributing factors, such as income.
+The first data explored is income data across the four designated regions in the United States (Northeast, Midwest, South, West).
+The adjusted incomes (per persons and per bedroom) are also considered, as certain regions may have different averages for household and property sizes, which could skew the unadjusted income data.
+The behavior of household income, bracketed in ranges in accordance to census data, over time is displayed to provide a better understanding of the changes in burden over time.
+The household incomes are then compared versus burden over time.
+here it is found that the ghiher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
+Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
+finally, to elucidate the gap in burden lvels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
+  A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
+This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
+
+###Graphical findings
+
+We start off by measuring the housing burden levels across the United States:
+
+```{r, echo = F}
+col_types <- "cicciiciiciiidicidicddddddiidddddddddddddddddididididididididididididididicccccccccccccccccccccccccd"
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
 
@@
     scale_colour_brewer(palette = "Dark2") +
-  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
     theme_tufte()
@@
 
-The graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
+  This graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
 The median burden varies dramatically between households of different sizes.
@@
 This is understandable, since people living alone cannot split any costs between other earners.
-
   Surprisingly, households of two people have the lowest burdens overall.
 Our hypothesis is that this is due to a large proportion of two-earner couples in this group.
-As housing cost burden is simply a household’s monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
+As housing cost burden is simply a household's monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
 In households of 3 or more people, the burden picks up again.
 This is likely due to a larger amount of children and / or retirees in larger households.
-
   Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007.
@@
 
-```{r, include = F, fig.cap="Median Income across Census Regions"}
+```{r, include = F}
 url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
@@
 median_incomes <- clean_years %>%
-  mutate(region = as.character(region)) %>%
   group_by(region) %>%
@@
                                  direction = 1,
-                                 guide = "legend",
-                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})
+                                 guide = "legend")
 ```
@@
 
-```{r, echo = FALSE, fig.cap="Amount of households relative to poverty line"}
+```{r, echo = FALSE}
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
 clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)", 
+       title = "Median Income Adjusted \nfor the Number of People by Region") + 
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte()
+
+```
+
+  This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+APLmed is the adjusted median household income per person.
+This was precaculated using Housing and Urban Development (HUD) indexes.
+For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
+The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+  It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+  The median adjusted income increases steadily (roughly linearly) over time.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
+economy) while the other regions took some time to correct for the downturn in the financial market.
+
+```{r, echo = F}
+
+clean_years <- read_csv("../data/clean_years.csv", col_types = col_types)
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) +
+  labs(x = "Year", y = "Median Adjusted Income (per bedroom)", 
+       title = "Median Income Adjusted \nfor Number of Bedrooms by region") + 
+  scale_colour_brewer(palette = "Dark2",
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_tufte()
+
+```
+
+  The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+ABlmed was calculated using precaculated HUD indexes.
+Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
+Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
+  It is interesting to note that the disparity between the incomes increases between the Northeast and WEstern regions and the Midwest and South.
+This is liekly because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
+This is not shown in the graphs, which indicate that the income levels are significantly different.
+However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
+  The lag in the downward trend after 2008 still occurs in the plots.
+
+```{r, echo = FALSE}
+counts <- clean_years %>% 
   # Group and summarise the amount of people in each poverty group each year
@@
                                              "> 200% of poverty income"),
-                                  ordered = T)) %>% 
-  ggplot(aes(x = year, 
+                                  ordered = T))
+
+ggplot(counts, aes(x = year, 
                    y = percent,
@@
                     guide = guide_legend(reverse=TRUE)) +
-  labs(x = "Year",
-       y = "Percent of households",
+  labs(title = "Amount of households relative to poverty line",
+       x = "Year",
+       y = "Precent of households",
        fill = "Household income
 relative to poverty income") +
-  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
   This plot shows the amount of households in a given income group in a given year.
-Each household has been calculated a poverty treshold according to the Census Bureau’s official thresholds, based on family size and number of children.
-
+Each household has been calculated a poverty treshold according to the Census Bureau's official thresholds, based on family size and number of children.
   As can be seen from the plot, the amount of people in different income groups stayed relatively constant between 1985 and 2007.
@@
 
-```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
+```{r, echo = FALSE}
+
+cleanyears <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
+  select(zinc2, year, region, fmtburden)%>% 
+  filter(fmtburden != ".") %>% 
+  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+
 ###### Household Income
   ####overall graph
-clean_years %>% 
-  select(zinc2, year, region, fmtburden)%>% 
-  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
+cleanyears %>%  
     group_by_("year", "fmtburden") %>%  
@@
     paste0("~ median(","zinc2", ",na.rm=T)"))) %>%
+
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
   geom_line(size = 1) + 
-  labs(y = "Household Income", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
+  labs(title = "Figure : Relationship between Household \nIncome and Burden", 
+       y = "Household Income", x = "Year", color = "Burden") +
+  scale_y_continuous(labels = unit_format(unit = "K")) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
+```{r, echo = FALSE}
+
  ### faceted by region 
-clean_years %>%  
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>%
-  filter(fmtburden != "4 No Income") %>% 
+  cleanyears %>%  
   group_by_("year","region", "fmtburden") %>%  
@@
     paste0("~ median(","zinc2", ",na.rm=T)"))) %>%
+
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
@@
   facet_wrap(~region, labeller = as_labeller(regionNames)) +
-  labs(y = "Household Income ($)", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,25000,50000,75000),labels=function(x){paste0(x/1000, "K")}) +
+  labs(title = "Relationship between Household \nIncome and Burden", 
+       y = "Household Income ($)", x = "Year", color = "Burden") +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo = FALSE, fig.cap = "Relationship between Monthly Housing Costs and Burden"}
+```{r, echo = FALSE}
 
-cleanyearsmonthly <- clean_years %>% 
+cleanyearsmonthly <- read_csv("../data/clean_years.csv", col_types = col_types) %>% 
   select(year, zsmhc, region, fmtburden) %>% 
@@
 ####overall graph
+
 cleanyearsmonthly %>%
+
   group_by_("year", "fmtburden") %>%  
@@
   geom_line(size = 1) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo= FALSE, fig.cap="Relationship between Monthly Housing Costs and Burden"}
+```{r, echo= FALSE}
 
@@
   cleanyearsmonthly %>%
+
   group_by_("year","region", "fmtburden") %>%  
@@
     paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%
+
   ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
@@
   facet_wrap(~region, labeller = as_labeller(regionNames)) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(title = "Relationship between Monthly \nHousing Costs and Burden", 
+       y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE, fig.cap = "Income compared to area's median grouped by house type"}
-clean_years %>% 
-  filter(fmtstructuretype != ".") %>% 
+```{r, fig.width=12, fig.height=6,echo = F, message=F,error=F, warning=FALSE}
+
+clean_sample <- read_csv("../data/clean_years.csv")
+clean_sample %>% filter(fmtstructuretype != ".") %>% 
       # Use NSE
@@
   geom_line(size = 1) + 
-  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
-  labs(colour = "Structure Type",
+  labs(title="Income compared to area's median grouped by house type",
+       colour = "Structure Type",
        y="Income percentage of area median income",
@@
 
-```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median housing cost with assistance"}
-assisted = clean_years %>% filter(!is.na(assisted)) %>% 
+```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE}
+assisted = clean_sample %>% filter(!is.na(assisted)) %>% 
       # Use NSE
@@
            stat="identity") + 
-  labs(y="Burden",
+  labs(title="Burden based on Median housing cost with assistance",
+       y="Burden",
        x = "",
        fill="Assistance") + 
-  scale_y_continuous(limits=c(0,1), labels=function(x){paste0(x*100, "%")}) +
-  scale_x_discrete(labels = function(x){str_replace_all(str_replace_all(x, "^[0-9] (.*)","\\1"), "LTE (.*)","<= \\1")}) +
+  scale_y_continuous(limits=c(0,1)) +
   scale_fill_brewer("Assistance", 
@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region"}
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(aplmed, na.rm = T)) %>%
-  mutate(region=as.character(region)) %>% 
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
-  scale_colour_brewer(palette = "Dark2", 
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_tufte()
+##Conclusion:
 
-```
-
-This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-It is clear that the Northeast and the West are the most expensive regions.
-This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
-
-The median adjusted income increases steadily (roughly linearly) over time.
-There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
-So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
-economy) while the other regions took some time to correct for the downturn in the financial market.
-
-```{r, echo = F, fig.cap="Median Income Adjusted for Number of Bedrooms by region"}
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(ablmed, na.rm = T)) %>%
-  mutate(region = as.character(region)) %>% 
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") +
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
-  scale_colour_brewer(palette = "Dark2",
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_tufte()
-
-```
-
-The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
-
-It is interesting to note that the disparity between the incomes increases between the Northeast/Western regions and the Midwest/Southern Regions.
-This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
-However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
-This is not shown in the graphs, which indicate that the income levels are significantly different.
-However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
-
-The lag in the downward trend after 2008 still occurs in the plots.
-
-Conclusion:
-
-Blah blah blah blah blah
+  The findings in this paper help elucidate income and househodl burden trends overtime.
+However, there are many holes in the actual data.
+The data grabbed uses gross incomes and groups data over large regions of the United States.
+While the spreads of incomes and housing costs may be low, the internal varation within each region will be fairly high.
+As a result, this does not help protray a full picture of burden level across the United States, namely in (sub)urban versus non urban areas.
+Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.

### 2016-08-08 paper/Paper.Rmd
edit text , reogranize

--- 2016-08-08 paper/Paper.Rmd
@@
 ---
-title: "Paper TITLE"
+title: "Exploring United States Housing Affordability"
 output: 
@@
 
-Summary/Intro
+##Introduction
 
-We chose the dataset from the American Housing Survey, which derives sets of data from 1985 and on.
+  The dataset was taken from the American Housing Survey, which derives sets of data from 1985 and on.
 The data set includes categorizing each housing unit by factors such as affordability, income through Fair Market Rent and the Adjusted Median Income.
@@
 Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+  The goal is to determine whether housing affordability has increased or decreased over the last few decades.
+Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
+The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
+Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
+Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
-We chose to analyze the data and ultimately view determines whether housing affordability has increased or decreased over the last few decades.
-Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
-The United State’s growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
-Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
-Between all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the housing market affordability status.
+#Abstract
 
-Cleaning data
+###Cleaning data
 
-Cleaning data
   Initially, the dataset is split into fifteen datasets, one for each odd year between 1985 and 2013.
@@
 Finally we combine every list of datasets into a single dataset, and then store it in a comma separated file called combined_years.
+  The combined_years data was read into a dataframe using the read_csv function.
+Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
+Also, the nonsensical integers (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
+Then, the combined years dataset was cleaned using the piping methods in dplyr.
+The data was rather messy, with unecessary quotations around character values and inconsistent letter cases in the column names.
+This was rectified by setting all the column names to lowercase and removing all the quotations from character values.
+One the main premises of this paper is comparing the burden of household that are owning and renting.
+Since there was no explicit labeling of data corresponding to owning and renting households, an  additional column "ownrent" was created which labeled levels 1 and 2 to own and rent as factors.
+Finally, the data was written to a clean_years csv file.
 
-The "combined_ years" data was read into a dataframe using the "read_ csv" function.
-Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), and the nonsensical integer date (negative numbers) were converted to NAs in the dataset.
-Then, the combined years dataset was cleaned using the piping methods in dplyr.
-All the column names were set to lowercase, and all the quotations were removed from character values.
-Then, an additional column “ownrent” was added which labeled levels 1 and 2 to own and rent as factors. Then, the data was written to a clean_years csv file.
+###Analyzing the Data
 
+  The main metric that is referred to in this paper is burden.
+Burden is defined by the proportion of household income that is spent towards the cost of housing.
+This metric is then further dissected by analyzing against various variables and collating data on contributing factors, such as income.
+The first data explored is income data across the four designated regions in the United States (Northeast, Midwest, South, West).
+The adjusted incomes (per persons and per bedroom) are also considered, as certain regions may have different averages for household and property sizes, which could skew the unadjusted income data.
+The behavior of household income, bracketed in ranges in accordance to census data, over time is displayed to provide a better understanding of the changes in burden over time.
+The household incomes are then compared versus burden over time.
+here it is found that the ghiher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
+Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
+finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
+  A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
+In particular, the lag in the downward spike after 2007 between the Northeast and the other regions is discussed.
+This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
+
+###Graphical findings
+
+  The inital plot measures the housing burden levels across the United States:
 ```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
@@
 This is understandable, since people living alone cannot split any costs between other earners.
-
   Surprisingly, households of two people have the lowest burdens overall.
 Our hypothesis is that this is due to a large proportion of two-earner couples in this group.
-As housing cost burden is simply a household’s monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
+As housing cost burden is simply a household's monthly housing cost divided by its monthly income, households where all members are earners have a higher monthly income and a lower resulting burden.
 In households of 3 or more people, the burden picks up again.
 This is likely due to a larger amount of children and / or retirees in larger households.
-
+Also, households with 3 or more members tend to hae only one earner.
   Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007.
@@
 We used median incomes instead of mean incomes to avoid skewing the results with outliers.
-
   The household median monthly incomes vary between different census regions from 32200 dollars in the South to nearly 25% higher, 40000, in the West.
 The Northeast trails closely behind the West with 37267.50 dollars, and the Midwest is closer to the South with a median income of 35000.
-
   While the dataset doesn't have more granular location data, even these four regions are very distinct from each other.
 The large high-income cities in the Northeast and the West likely contribute to the higher respective median incomes, while more rural Midwest and South are poorer.
+
+```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=5, fig.height=5}
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(aplmed, na.rm = T)) %>%
+  mutate(region=as.character(region)) %>% 
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) + 
+  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
+  scale_colour_brewer(palette = "Dark2", 
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_minimal(base_family = "serif")
+
+clean_years %>%
+  group_by(year, region) %>%  
+  # Take the median instead of mean to avoid extreme observations skewing the 
+  # result.
+  summarise(new_y = median(ablmed, na.rm = T)) %>%
+  mutate(region = as.character(region)) %>% 
+  ggplot(aes(x = year,   
+                  y = new_y,   
+                  colour = region)) + 
+  geom_line(size = 1) +
+  labs(x = "Year", y = "Median Adjusted Income (per bedroom)") +
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
+  scale_colour_brewer(palette = "Dark2",
+                      name = "Region", 
+                      labels = regionNames) +
+  theme_minimal(base_family = "serif")
+
+```
+
+  In order to better understand the regional differences in income, the incomes were adjusted by certain variables. 
+  The first plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+APLmed is the adjusted median household income per person.
+This was precaculated using Housing and Urban Development (HUD) indexes.
+For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
+The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+  It is clear that the Northeast and the West are the most expensive regions.
+This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+  The median adjusted income increases steadily (roughly linearly) over time.
+There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
+economy) while the other regions took some time to correct for the downturn in the financial market.
+  The second plot of the graph depicts ABLmed versus year.
+ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
+ABlmed was calculated using precaculated HUD indexes.
+Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
+Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
+  It is interesting to note that the disparity between the incomes increases between the Northeast and Western regions and the Midwest and South.
+This is likely because those living in the Midwest and South tend to live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
+However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
+This is not shown in the graphs, which indicate that the income levels are significantly different.
+However, if housing affordability was factored into the calculation, the income disparity would be far less if not almost nonexistent.
+The lag in the downward trend after 2008 still occurs in the plots.
+  With the income information in hand, the burden data can be explored in further detail.
 
@@
 
+ Further delving into the behavior of burden amongst the lower income workers, this plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
+At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance.
+This clearly makes sense, as those who need assistance are typically people who've lost their job or had some monetary crisis that would reduce their total salary.
+However, although assistance has around the same effect for most cost brackets, the households that have excessively high costs for their area have vastly higher burdens with assistance than all other brackets.
+One explanation for this is that those who purchase expensive houses in their area typically have higher salaries, and if they lose those salaries and require assistance they have to support the cost of the house.
+Also, single family homeowners have significantly less mobility so if they have a financial crisis they cannot easily transition to cheaper housing as opposing to those with rental properties typical of a larger number of units.
 The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=5, fig.height=5}
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(aplmed, na.rm = T)) %>%
-  mutate(region=as.character(region)) %>% 
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
-  scale_colour_brewer(palette = "Dark2", 
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_minimal(base_family = "serif")
+##Conclusion:
 
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(ablmed, na.rm = T)) %>%
-  mutate(region = as.character(region)) %>% 
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") +
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
-  scale_colour_brewer(palette = "Dark2",
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_minimal(base_family = "serif")
-
-```
-
-This plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
-It is clear that the Northeast and the West are the most expensive regions.
-This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
-
-The median adjusted income increases steadily (roughly linearly) over time.
-There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
-So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
-economy) while the other regions took some time to correct for the downturn in the financial market.
-
-```{r, echo = F, fig.cap="Median Income Adjusted for Number of Bedrooms by region"}
-clean_years %>%
-  group_by(year, region) %>%  
-  # Take the median instead of mean to avoid extreme observations skewing the 
-  # result.
-  summarise(new_y = median(ablmed, na.rm = T)) %>%
-  mutate(region = as.character(region)) %>% 
-  ggplot(aes(x = year,   
-                  y = new_y,   
-                  colour = region)) + 
-  geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") +
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
-  scale_colour_brewer(palette = "Dark2",
-                      name = "Region", 
-                      labels = regionNames) +
-  theme_minimal(base_family = "serif")
-
-```
-
-The trends of the graphs here are similar to the ones in APLmed, however ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
-
-It is interesting to note that the disparity between the incomes increases between the Northeast/Western regions and the Midwest/Southern Regions.
-This is liekly because those living in the Midwest and South ten dto live in larger homes, housing more bedrooms on average, relative to their incomes due to the higher housing affordability.
-However, this should mean that their incomes go 'further,' as they have more purchasing parity given the lower cost of living compared to the Northeast and the West.
-This is not shown in the graphs, which indicate that the income levels are significantly different.
-However, if housing affordability was factored into the calcualtion, the income disparity would be far less if not almost nonexistent.
-
-The lag in the downward trend after 2008 still occurs in the plots.
-
-Conclusion:
-
-Blah blah blah blah blah
+  The findings in this paper help elucidate income and househodl burden trends overtime.
+However, there are many holes in the actual data.
+The data grabbed uses gross incomes and groups data over large regions of the United States.
+While the spreads of incomes and housing costs may be low, the internal varation within each region will be fairly high.
+As a result, this does not help protray a full picture of burden level across the United States, namely in (sub)urban versus non urban areas.
+Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
+  Further analysis could involve predicting the trends for the burden and income data to corroborate whether or not a two year lag exists between the Northeast and the other regions following the financial meltdown.
+Also, adjusting income by cost of living would provide a better understanding of burden amongst differing regions.
+Regions could be further subsetted by town or district level to provide more granular data.

### 2016-08-09 paper/Paper.Rmd
edit spelling
Switch the position of a variable definition

@@
   The inital plot measures the housing burden levels across the United States:
+
 ```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
-clean_years <- read_csv("../clean_data/clean_years.csv")
+clean_years <- read_csv("../clean_data/clean_sample.csv")
 
@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=5, fig.height=5}
+```{r, echo = FALSE}
+regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
+
+```
+
+```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=4, fig.height=4}
 clean_years %>%
@@
   geom_line(size = 1) + 
-  labs(x = "Year", y = "Median Adjusted Income (per persons)") + 
+  labs(x = "Year", title = "Median Adjusted Income (per persons)") + 
   scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
@@
   geom_line(size = 1) +
-  labs(x = "Year", y = "Median Adjusted Income (per bedroom)") +
+  labs(x = "Year", title = "Median Adjusted Income (per bedroom)") +
   scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
@@
 The reality can be more nuanced, with people dropping to the middle income groups and a simultaneous equally large movement from these brackets to poverty, but the big picture of large impacts in poverty following the financial crisis stays the same.
-
-```{r, echo = FALSE}
-regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
-
-```
 
@@
 
-  The findings in this paper help elucidate income and househodl burden trends overtime.
+  The findings in this paper help elucidate income and household burden trends overtime.
 However, there are many holes in the actual data.

### 2016-08-09 paper/Paper.Rmd
spaces

@@
 Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+
   The goal is to determine whether housing affordability has increased or decreased over the last few decades.
@@
 Finally we combine every list of datasets into a single dataset, and then store it in a comma separated file called combined_years.
+
   The combined_years data was read into a dataframe using the read_csv function.
@@
 Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
+
   A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
@@
 This is understandable, since people living alone cannot split any costs between other earners.
+
   Surprisingly, households of two people have the lowest burdens overall.
@@
 Also, households with 3 or more members tend to hae only one earner.
+
   Another interesting thing is the dramatic increase of burden during the buildup to the financial crisis 2000-2007.
@@
 We used median incomes instead of mean incomes to avoid skewing the results with outliers.
+
   The household median monthly incomes vary between different census regions from 32200 dollars in the South to nearly 25% higher, 40000, in the West.
 The Northeast trails closely behind the West with 37267.50 dollars, and the Midwest is closer to the South with a median income of 35000.
+
   While the dataset doesn't have more granular location data, even these four regions are very distinct from each other.
@@
   In order to better understand the regional differences in income, the incomes were adjusted by certain variables.
+
   The first plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
@@
 The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+
   It is clear that the Northeast and the West are the most expensive regions.
 This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+
   The median adjusted income increases steadily (roughly linearly) over time.
@@
 economy) while the other regions took some time to correct for the downturn in the financial market.
+
   The second plot of the graph depicts ABLmed versus year.
@@
 Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
-Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
+Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms.
+
   It is interesting to note that the disparity between the incomes increases between the Northeast and Western regions and the Midwest and South.
@@
 The lag in the downward trend after 2008 still occurs in the plots.
+
   With the income information in hand, the burden data can be explored in further detail.
@@
 Each household has been calculated a poverty treshold according to the Census Bureau’s official thresholds, based on family size and number of children.
+
   As can be seen from the plot, the amount of people in different income groups stayed relatively constant between 1985 and 2007.
@@
 This stable relationship broke noticeably in the financial crisis of 2007 and the following recession, when the amount of people in poverty started a quite steep and steady climb upwards.
+
   Interestingly, the income brackets of 100%-200% stayed almost exactly the same size as before, while the increase in poverty was in effect solely the result of the dwindling of the highest (over 200% of poverty income) income bracket.
@@
 Also, single family homeowners have significantly less mobility so if they have a financial crisis they cannot easily transition to cheaper housing as opposing to those with rental properties typical of a larger number of units.
+
   The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
@@
 Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
+
   Further analysis could involve predicting the trends for the burden and income data to corroborate whether or not a two year lag exists between the Northeast and the other regions following the financial meltdown.

### 2016-08-09 paper/Paper.Rmd
Fixing plots

--- 2016-08-09 paper/Paper.Rmd
@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=4, fig.height=4}
+```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=3.5, fig.height=4}
 clean_years %>%
@@
                       labels = regionNames) +
-  theme_minimal(base_family = "serif")
+  theme_minimal(base_family = "serif") +
+  theme(legend.position = "none", axis.title.y = element.blank())
 

### 2016-08-09 paper/Paper.Rmd

@@
   theme_minimal(base_family = "serif") +
-  theme(legend.position = "none", axis.title.y = element.blank())
+  theme(legend.position = "none")
 

### 2016-08-09 paper/Paper.Rmd
Fix side-by-side figure

@@
     fig_caption: yes
-    latex_engine: xelatex
+    latex_engine: 
+header-includes:
+#- \usepackage{subfig}
+- \usepackage{subcaption}
 ---
@@
 
-```{r, echo=F}
+```{r, echo=F, fig.cap="Median Income across Census Regions"}
 usa_map
@@
 
-```{r, echo = FALSE, fig.cap="Median Income Adjusted for the Number of People by Region", fig.width=3.5, fig.height=4}
+```{r, echo = FALSE, fig.height=1, fig.cap='Median Adjusted Income per person (left) and per bedroom (right)'}
+ggplot() + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
+```
+
+```{r, echo = FALSE, fig.width=3.5, fig.height=5, fig.cap='one plotthe other one'}
 clean_years %>%
@@
   geom_line(size = 1) + 
-  labs(x = "Year", title = "Median Adjusted Income (per persons)") + 
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
+  labs(x = "Year") + 
+  scale_y_continuous(labels = function(x){paste0(x/1000, "K")},
+                     limits = c(20000, 80000)) +
   scale_colour_brewer(palette = "Dark2", 
@@
                       labels = regionNames) +
-  theme_minimal(base_family = "serif") +
-  theme(legend.position = "none")
+  theme_minimal(base_size = 10,
+                base_family = "serif") +
+  theme(legend.position = "top",
+        axis.title.y = element_blank())
 
@@
   geom_line(size = 1) +
-  labs(x = "Year", title = "Median Adjusted Income (per bedroom)") +
-  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")}) +
+  labs(x = "Year") +
+  scale_y_continuous(labels=function(x){paste0(x/1000, "K ")},
+                     limits = c(20000, 80000)) +
   scale_colour_brewer(palette = "Dark2",
@@
                       labels = regionNames) +
-  theme_minimal(base_family = "serif")
-
+  theme_minimal(base_size = 10,
+                base_family = "serif") +
+  theme(axis.title.y = element_blank(), 
+        axis.text.y = element_blank(),
+        legend.position = "top")
 ```
@@
 This was precaculated using Housing and Urban Development (HUD) indexes.
-For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household beig adjusted by 1.00 (or having no change from the median household income).
-The reason for no adjustment to a four person househodl income is becasue the median hosuehold size in the United States is about four persons.
+For example, a one person household's income is adjusted by a factor of 0.70, two by 0.80, and so on with a four person household being adjusted by 1.00 (or having no change from the median household income).
+The reason for no adjustment to a four person household income is because the median household size in the United States is about four persons.
   It is clear that the Northeast and the West are the most expensive regions.

### 2016-08-09 paper/Paper.Rmd
editing graphs and text

@@
     geom_line(size = 1) +
-    labs(colour = "Household size",
+    labs(colour = "Household Size",
          x = "Year",
-         y = "Median burden") +
+         y = "Median Burden") +
     scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, include = F, fig.cap="Median Income across Census Regions"}
+```{r, include = F, fig.cap = "Median Income Across Census Regions"}
 url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
@@
 median_incomes <- clean_years %>%
+  filter(year == 2013) %>%
   mutate(region = as.character(region)) %>%
@@
 
-usa_map <- map_df %>% 
+us_map %>% 
   ggplot() +  
@@
   coord_fixed(1.3) +
-  theme_minimal(base_family = "serif") +
+  theme_tufte(base_size = 20, base_family = "serif") +
   scale_x_continuous(name = NULL,
@@
                                  guide = "legend",
-                        labels=function(x){paste0("$",x %/% 100 / 10, "K")})
-```
-
-```{r, echo=F, fig.cap="Median Income across Census Regions"}
-usa_map
+                        labels=function(x){paste0("$",
+                                                  x %/% 100 / 10, "K")})
 ```
@@
 ```{r, echo = FALSE}
+
 regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
@@
 ```{r, echo = FALSE, fig.height=1, fig.cap='Median Adjusted Income per person (left) and per bedroom (right)'}
+
 ggplot() + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
@@
 The reason for no adjustment to a four person household income is because the median household size in the United States is about four persons.
+
   It is clear that the Northeast and the West are the most expensive regions.
 This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
+
   The median adjusted income increases steadily (roughly linearly) over time.
@@
 Three bedroom properties were adjusted by a factor of 1.04, because the median number of bedrooms in a property is roughly three bedrooms
+
   It is interesting to note that the disparity between the incomes increases between the Northeast and Western regions and the Midwest and South.
@@
 
-```{r, echo = FALSE, fig.cap="Amount of households relative to poverty line"}
+```{r, echo = FALSE, fig.cap = "Amount of households relative to poverty line"}
 clean_years %>% 
@@
   labs(x = "Year",
-       y = "Percent of households",
-       fill = "Household income
-relative to poverty income") +
+       y = "Percent of Households",
+       fill = "Household Income
+Relative to Poverty Income") +
   scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
@@
 clean_years %>% 
-  select(zinc2, year, region, fmtburden)%>% 
-  filter(fmtburden != ".", fmtburden != "4 No Income") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>% 
+  select(zinc2, 
+         year, 
+         region, 
+         fmtburden)%>% 
+  filter(fmtburden != ".", 
+         fmtburden != "No Income") %>% 
+  mutate(fmtburden = str_replace(fmtburden, 
+                                 "Greater than 50%", 
+                                 "50% or More")) %>% 
   group_by_("year", "fmtburden") %>%  
@@
   paste0("~ median(","zinc2", ",na.rm=T)"))) %>% 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  ggplot(aes_string(x = "year", 
+                    y = "new_y", 
+                    color = "fmtburden")) + 
   geom_line(size = 1) + 
-  labs(y = "Household Income", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,20000,40000,60000),labels=function(x){paste0(x/1000, "K")}) +
+  labs(y = "Household Income", 
+       x = "Year", 
+       color = "Burden") +
+  scale_y_continuous(limits = c(0,NA), 
+                     breaks=c(0,20000,40000,60000),
+                     labels=function(x){paste0("$",x/1000, "K")}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
- ### faceted by region 
-clean_years %>%  
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More")) %>%
-  filter(fmtburden != "4 No Income") %>% 
-  group_by_("year","region", "fmtburden") %>%  
-  summarise_(new_y = as.formula(
-    paste0("~ median(","zinc2", ",na.rm=T)"))) %>% 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
-  geom_line(size = 1) + 
-  facet_wrap(~region, labeller = as_labeller(regionNames)) +
-  labs(y = "Household Income ($)", x = "Year", color = "Burden") +
-  scale_y_continuous(limits=c(0,NA),breaks=c(0,25000,50000,75000),labels=function(x){paste0(x/1000, "K")}) +
-  scale_colour_brewer(palette = "Dark2") +
-  theme_minimal(base_family = "serif")
-
-```
-
-  From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
-
 ```{r, echo = FALSE, fig.cap = "Relationship between Monthly Housing Costs and Burden"}
@@
   filter(fmtburden != ".") %>% 
-  mutate(fmtburden = str_replace(fmtburden, "Greater than 50%", "50% or More"))
+  mutate(fmtburden = str_replace(fmtburden, 
+                                 "Greater than 50%", 
+                                 "50% or More"))
 
-####overall graph
 cleanyearsmonthly %>%  
@@
   summarise_(new_y = as.formula(
-    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>%
+    paste0("~ median(","zsmhc", ", na.rm=T)"))) %>%
 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+  ggplot(aes_string(x = "year", 
+                    y = "new_y", 
+                    color = "fmtburden")) + 
   geom_line(size = 1) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(y = "Monthly Housing Costs", 
+       x = "Year", 
+       color = "Burden") +
+  scale_y_continuous(labels=function(x){paste0("$", x)}) +
   scale_colour_brewer(palette = "Dark2") +
@@
   facet_wrap(~region, labeller = as_labeller(regionNames)) + 
-  labs(y = "Monthly Housing Costs ($)", x = "Year", color = "Burden") +
+  labs(y = "Monthly Housing Costs", x = "Year", color = "Burden") +
+  scale_y_continuous(labels=function(x){paste0("$", x)}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
+  From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
 From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
@@
   geom_line(size = 1) +
-  scale_y_continuous(labels=function(x){paste0(x*100, "%")}) +
+  scale_y_continuous(labels = function(x){paste0(x, "%")}) +
   labs(colour = "Structure Type",
-       y="Income percentage of area median income",
+       y = "Income Percentage of Area Median Income",
        x = "Year") +

### 2016-08-10 paper/Paper.Rmd
grammer and pushing

@@
 ```
+#Abstract
 
@@
   The dataset was taken from the American Housing Survey, which derives sets of data from 1985 and on.
-The data set includes categorizing each housing unit by factors such as affordability, income through Fair Market Rent and the Adjusted Median Income.
-Taking into account variables such as Burden (housing cost divided by its monthly income), certain households with a zero or negative income are input as -1 within the data.
-The data source was recorded from the American Housing Survey.
+The data set includes categorizing each housing unit by factors such as affordability, income through fair market rent, burden, and the adjusted median income.
 Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
@@
 Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
-
-#Abstract
 
@@
   The combined_years data was read into a dataframe using the read_csv function.
-Within the read_csv function, the column types were explicitly stated as a string variable (c = character, i = integer, etc.), in order for the data to be wrangled properly.
-Also, the nonsensical integers (negative numbers) were converted to NAs in the dataset so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
 Then, the combined years dataset was cleaned using the piping methods in dplyr.
@@
 One the main premises of this paper is comparing the burden of household that are owning and renting.
-Since there was no explicit labeling of data corresponding to owning and renting households, an  additional column "ownrent" was created which labeled levels 1 and 2 to own and rent as factors.
+Since there was no explicit labeling of data corresponding to owning and renting households, the "ownrent" column was mutated to a factor with the correct labels.
 Finally, the data was written to a clean_years csv file.
+The source files used different nonsensical integers (negative numbers) to indicate NA values.
+These were converted to NAs while reading the dataset in so that the output data and graphs made stastical sense - negative numbers where the values should be positive skew the dataset.
 
@@
 The household incomes are then compared versus burden over time.
-here it is found that the ghiher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
+here it is found that the higher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
 Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
-finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
-Here it is found thatowning households tend to have lower burden levels, aso do those households with no assistance.
+Finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+Here it is found that owning households tend to have lower burden levels, as do those households with no assistance.
 
@@
 
-  The inital plot measures the housing burden levels across the United States:
-
 ```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
-clean_years <- read_csv("../clean_data/clean_sample.csv")
+clean_years <- read_csv("../clean_data/clean_years.csv")
 
@@
 
-  This graph shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
+  Figure 1 shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
 The median burden varies dramatically between households of different sizes.
@@
 
-```{r, include = F, fig.cap = "Median Income Across Census Regions"}
+```{r, echo = F, fig.cap = "Median Income Across Census Regions", message = F, warning=F}
 url <- "http://www.mapsofworld.com/usa/usa-maps/united-states-regional-maps.html"
@@
 
-  To gauge income levels in different areas we used the four census regions in the dataset and joined them with a scraped table from mapsoftheworld.com.
+  To map income levels in different areas we used the four census regions in the dataset and joined them with a scraped table from mapsoftheworld.com.
 Using the table we could connect census regions to state names and thus colour a map of the continental US with census region incomes.
-We used median incomes instead of mean incomes to avoid skewing the results with outliers.
+We used median incomes instead of mean incomes to avoid skewing the results with outliers, and filtered the map to include the most recent year, 2013.
 
-  The household median monthly incomes vary between different census regions from 32200 dollars in the South to nearly 25% higher, 40000, in the West.
-The Northeast trails closely behind the West with 37267.50 dollars, and the Midwest is closer to the South with a median income of 35000.
+  The resulting figure 2 shows household median monthly incomes vary between different census regions from $39,900 in the South to nearly 25% higher, $49,600, in the Northeast.
+The West trails closely behind the Northeast with $48,000, and the Midwest is closer to the South with a median income of $43,900.
 
@@
 ```
+
+\newpage
 
@@
 
+Figure 3: Median Adjusted Income per person (left) and per bedroom (right)
+
   In order to better understand the regional differences in income, the incomes were adjusted by certain variables.
 
-  The first plot compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+  The first plot of figure 3 compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
 APLmed is the adjusted median household income per person.
@@
 There is a decrease in income around 2010, correlated to the economic recession with a 2 year lag between the Northeast and the rest of the regions.The reason for this lag may be due to the fact that the financial center of the United States lies in the northeast, around New York City.
-So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street and Main Street- which refers to the Washington D.C.
-economy) while the other regions took some time to correct for the downturn in the financial market.
+So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street) while the other regions took some time to correct for the downturn in the financial market.
 
-  The second plot of the graph depicts ABLmed versus year.
-ABLmed signifies the median adjusted income for the number of bedrooms in a household's living quarters.
-ABlmed was calculated using precaculated HUD indexes.
+  The second plot of figure 3 depicts the development of median income adjusted for the number of bedrooms in a household's living quarters.
+The adjusted median income was calculated using precaculated HUD indexes.
 Here, zero bedrooms was adjusted by a factor of 0.70, two bedrooms by 0.90, and so on.
@@
 
-  This plot shows the amount of households in a given income group in a given year.
+  Figure 4 shows the amount of households in a given income group in a given year.
 Each household has been calculated a poverty treshold according to the Census Bureau’s official thresholds, based on family size and number of children.
@@
 
-  As pictured in the graph, the relationship between household income and burden is seen as a positive correlation.
+  As pictured in figure 5, the relationship between household income and burden is seen as a positive correlation.
 From the graph, you can see that higher income household have less burden to deal with, which makes sense because those with more money are able to pay off monthly costs more easily.
 Those with 50% more burden are near the poverty line.
-From this graph, you can see a spike in households that have burden.
+Figure 5 also shows a spike in households that have burden.
 Households with higher incomes started to pay more burdens, which can be attributed to 2007 financial housing market crisis.
@@
 
-  From this graph, you can take a more indepth look at the relationship between housing costs and burden.
+  Figure 6 offers a more in-depth look at the relationship between housing costs and burden.
 It's interesting to see that all three burden lines are close to one another, which illustrates the principle that those with more money aren't buying more expensive houses.
-From the previous grpah, we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
+From figure 5 we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
 Each level of burden has similar housing costs, therefore reinforcing the idea that people are buying similar priced housing among the lower and middle classes.
@@
 
-  From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
-From this graph, you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
+  From figure 7 you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
 
@@
 
-  This plot charts a time series showing how the typical income for various types of households such as single family or apartment buildings changed over the years.
+  Figure 8 charts a time series showing how the typical income for various types of households such as single family or apartment buildings changed over the years.
 One major trend is that typical income of all housing types except single family homes compared to their area has fallen, accelerating after 2007.
@@
 The first is that, with the growth of wages, more people could afford single family homes and thus leave the buildings with higher numbers of units.
-As wage grew, the area's median income also grew, and so, combined with higher waged workers moving into single family houses, single family homes remained constant while all other types shrunk proportionally.
+As wages increased, the area's median income also grew, and so, combined with higher waged workers moving into single family houses, single family homes remained constant while all other types shrunk proportionally.
 This would also explain the acceleration after the housing crisis, as single family homes became much cheaper temporarily and thus allow more workers to obtain single family housing.
@@
 
- Further delving into the behavior of burden amongst the lower income workers, this plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
-At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance.
+ Further delving into the behavior of burden amongst the lower income workers, figure 9 shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
+At initial observation, the burden of those who receive more assistance is higher than those who receive no assistance.
 This clearly makes sense, as those who need assistance are typically people who've lost their job or had some monetary crisis that would reduce their total salary.
@@
 The plot shows how burden changes as a result of assistance based on what percentage of the area median income is their monthly housing costs.
+
  At initial observation, the burden of those who recieve more assistance is higher than those who recieve no assistance.

### 2016-08-10 paper/Paper.Rmd
paper grammer

@@
 ```
+
+<br>
+<br>
+<br>
+
 #Abstract
-
-##Introduction
 
@@
 The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
-Another major event that could have potentially affected the current state of housing affordability, is the 2007 Housing Market crash that affected the entire country, and even the world.
+Another major event that has potential consequences on the current state of housing affordability, is the 2007 Housing Market crash.
 Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
-###Cleaning data
+The main metric that is referred to in this paper is burden.
+Burden is defined by the proportion of household income that is spent towards the cost of housing.
+This metric is then further dissected by analyzing against various variables and collating data on contributing factors, such as income.
+The first data explored is income data across the four designated regions in the United States (Northeast, Midwest, South, West).
+The adjusted incomes (per persons and per bedroom) are also considered, as certain regions may have different averages for household and property sizes, which could skew the unadjusted income data.
+The behavior of household income, bracketed in ranges in accordance to census data over time, is displayed to provide a better understanding of the changes in burden over time.
+The household incomes are then compared versus burden over time.
+It is found that the higher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
+Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
+Finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
+Here it is found that owning households tend to have lower burden levels, as do those households with no assistance.
+
+  A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
+In particular, the lag in the downward spike after 2007 between the Northeast and the other regions is discussed.
+This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
+
+  \newpage
+
+##Cleaning the Data
 
@@
 
-###Analyzing the Data
-
-  The main metric that is referred to in this paper is burden.
-Burden is defined by the proportion of household income that is spent towards the cost of housing.
-This metric is then further dissected by analyzing against various variables and collating data on contributing factors, such as income.
-The first data explored is income data across the four designated regions in the United States (Northeast, Midwest, South, West).
-The adjusted incomes (per persons and per bedroom) are also considered, as certain regions may have different averages for household and property sizes, which could skew the unadjusted income data.
-The behavior of household income, bracketed in ranges in accordance to census data, over time is displayed to provide a better understanding of the changes in burden over time.
-The household incomes are then compared versus burden over time.
-here it is found that the higher income levels generally have lower housing burdens, and those burdens actually decrased over time, while burdens in the lower income brackets were generally higher and increased over time.
-Also, the housing burdens in the Northeast and West are generally higher than the Midwest and South.
-Finally, to elucidate the gap in burden levels across high and low incomes, the housheold burden amongst owners and renters, and those with and without governement assistance is considered.
-Here it is found that owning households tend to have lower burden levels, as do those households with no assistance.
-
-  A common issue adressed throughout this paper is the anomaly in datasets around the 2007-2011 period, which  roughly marks the duration of the financial crisis.
-In particular, the lag in the downward spike after 2007 between the Northeast and the other regions is discussed.
-This section of the data helps in understading the dynamics of burden and income, as the financial crisis was recent and nationwide.
-
-###Graphical findings
+##Graphical findings
 
@@
 
-##Conclusion:
+#Conclusion:
 

### 2016-08-10 paper/Paper.Rmd
fixed the intro and abstract

@@
 
-#Abstract
+#Introduction
 
-The dataset was taken from the American Housing Survey, which derives sets of data from 1985 and on.
+The dataset was taken from the American Housing Survey (AHS), which derives sets of data from 1985 and on.
 The data set includes categorizing each housing unit by factors such as affordability, income through fair market rent, burden, and the adjusted median income.
@@
 Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
+
+#Abstract
 
@@
 Also, the usage of gross income is inaccurate, as the median net income levels varied over time (even after adjusting for inflation), since taxation levels were much higher in the 1950s and 1960s than they are now- which means the burden levels are likely undervalued in the data during those decades.
+Also, the data provided is given in odd years, which prevents a complete analysis of housing affordability.
+
+    Ultimately, our findings point towards
 

### 2016-08-10 paper/Paper.Rmd
edits

--- 2016-08-10 paper/Paper.Rmd
@@
 title: "Exploring United States Housing Affordability"
+author: "Nikolai Myllymäki,\nWilliam Paul,\nSriharsha Kondapally,\nLiam Shi"
 output: 
   pdf_document: 
-    fig_caption: yes
-    latex_engine: 
-header-includes:
-#- \usepackage{subfig}
-- \usepackage{subcaption}
+    fig_caption: no
+
 ---
@@
 
-```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
+```{r, echo = F, fig.cap="median burden across years and household sizes"}
 clean_years <- read_csv("../clean_data/clean_years.csv")
@@
 ```
+\begin{center}
+Figure 1: Median burden across years and household sizes
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 2: Median Income Across Census Regions
+\end{center}
 
@@
 
-```
-
-\newpage
-
-```{r, echo = FALSE, fig.height=1, fig.cap='Median Adjusted Income per person (left) and per bedroom (right)'}
-
-ggplot() + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
 ```
@@
 ```
-
+\begin{center}
 Figure 3: Median Adjusted Income per person (left) and per bedroom (right)
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 4: Amount of households relative to poverty line
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 5: Relationship between Household Income and Burden
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 6: Relationship between Monthly Housing Costs and Burden
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 7: Relationship between Monthly Housing Costs and Burden in different regions
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 8: Income compared to area's median grouped by house type
+\end{center}
 
@@
 ```
+\begin{center}
+Figure 9: Burden based on Median housing cost with assistance
+\end{center}
 

### 2016-08-10 paper/Paper.Rmd
minor changes

--- 2016-08-10 paper/Paper.Rmd
@@
 
-##Graphical findings
+#Graphical findings
 
 ```{r, echo = F, fig.cap="Median Burden across Years and Household sizes"}
+#Read in the cleaned data
 clean_years <- read_csv("../clean_data/clean_years.csv")
@@
 
-region_1 <- regions_tbl[2:7,] %>% 
-  gather(key = region, value = state, 1:2) %>% 
+region_1 <- regions_tbl[2:7, ] %>% 
+  gather(key = region, 
+         value = state, 1:2) %>% 
   slice(1:9) %>%
@@
 
-region_2 <- regions_tbl[10:15,] %>% 
-  gather(key = region, value = state, 1:2) %>% 
+region_2 <- regions_tbl[10:15, ] %>% 
+  gather(key = region, 
+         value = state, 1:2) %>% 
   mutate(region = 2)
@@
 region_3 <- regions_tbl[18:27,] %>%  
-  gather(key = region, value = state, 1:2) %>% 
+  gather(key = region, 
+         value = state, 1:2) %>% 
   slice(c(-10, -15, -16)) %>%
@@
   regions_tbl[30:37,] %>%  
-  gather(key = region, value = state, 1:2) %>% 
+  gather(key = region, 
+         value = state, 1:2) %>% 
   slice(1:13) %>%
@@
 
-regions <- rbind(region_1, region_2, region_3, region_4) %>%
+regions <- rbind(region_1, 
+                 region_2, 
+                 region_3, 
+                 region_4) %>%
   mutate(state = str_to_title(state),
@@
   coord_fixed(1.3) +
-  theme_tufte(base_size = 20, base_family = "serif") +
+  theme_tufte(base_size = 20, 
+              base_family = "serif") +
   scale_x_continuous(name = NULL,
@@
 
-regionNames = c("1" = "Northeast", "2" = "Midwest", "3" = "South", "4" = "West")
+#Factoring for the region faceting
+
+regionNames = c("1" = "Northeast", 
+                "2" = "Midwest", 
+                "3" = "South", 
+                "4" = "West")
 
@@
 
-ggplot() + theme(panel.background = element_rect(fill = 'white', colour = 'white'))
+ggplot() + 
+  theme(panel.background = element_rect(fill = 'white', 
+                                        colour = 'white'))
+
 ```

### 2016-08-10 paper/Paper.Rmd

--- 2016-08-10 paper/Paper.Rmd
@@
 
-##Graphical findings
+#Graphical findings
 
-```{r, echo = F, fig.cap="median burden across years and household sizes"}
+```{r, echo = F, fig.cap="Median Burden Across Years and Household Sizes"}
+#Read in the cleaned data
+
 clean_years <- read_csv("../clean_data/clean_years.csv")

### 2016-08-10 clean_data/CleanData.R
Added comments

@@ -41,29 +41,44 @@ RenameCols <- function(frame) {
          "METRO3" = `METRO`)
 }
 
-base_files <- paste("../raw_data/hads",          # Loading the most recent datasets 
-                c(seq(2003, 2013, by = 2)), # first to give reference for other
-                ".bz2",                     # two sets.
+NoQuotations <- function(x) {
+  # Removes quotes from input string x.
+  str_replace_all(x, "^\\'|\\'$", "")
+}
+# Total list of invalid values, done due to insufficient documentation.
+NA_LIST = c("","NA","-5","-6","-7","-9",".")
+# Loading the most recent datasets first to give reference for other
+# two sets.
+base_files <- paste("../raw_data/hads",          
+                c(seq(2003, 2013, by = 2)),  
+                ".bz2",                     
                 sep = "") %>%  
-  lapply(read_csv) %>% 
-  lapply(data.frame,  # Have to convert to data frames as lapply seems 
-         stringsAsFactors = F) #to recognize csv output as named lists.
-
-base_files <- mapply(MutateYears, # Need to distinguish different years
-                 base_files,     #for datasets, so add column to do just that.
+  lapply(read_csv, na=NA_LIST) %>% 
+  # Have to convert to data frames as lapply seems to recognize csv output 
+  # as named lists.
+  lapply(data.frame,   
+         stringsAsFactors = F)
+# Need to distinguish different years for datasets, so we add a column 
+# to do just that.
+base_files <- mapply(MutateYears, 
+                 base_files,     #
                  as.list(seq(2003, 2013, by = 2)))
 
-modified_year <- base_files[[4]] %>% # The 2009 dataset has an extra unspecified
-  select(-VCHRMOV)                  # column that we need to remove.
-modern_files <- rbind(do.call(rbind, # Combine all of the recent years into one.
+# The 2009 dataset has an extra unspecified column that we need to remove.
+modified_year <- base_files[[4]] %>% 
+  select(-VCHRMOV)      
+
+# Combine all of the recent years into one. 
+modern_files <- rbind(do.call(rbind, 
                             base_files[c(1, 2, 3, 5, 6)]), 
                     modified_year)
-
-first_decade <- paste("../raw_data/hads",             # Load first decade separately 
-                     c(seq(1985, 1995, by = 2)), # to account for missing 
[…]

### 2016-08-10 raw_data/download_dataset.R

@@ -1,0 +1,42 @@
+library(readr)
+library(stringr)
+# This script downloads the datasets for our project and reformats them 
+# into the correct form.
+
+downloadfile = function(url, filename, outfilename){
+  # This function downloads a file from the dataset website, unzips it, 
+  # and compresses it into a bz2.
+  #
+  # Args:
+  #  url: Name of zipped file
+  #  filename: Name of file inside zip
+  #  outfilename: Name of Output file
+  prefix_url = "https://www.huduser.gov/portal/datasets/hads/"
+  temp = tempfile()
+  download.file(str_c(prefix_url, url), temp)
+  csv = read_csv(unz(temp, filename))
+  unlink(temp)
+  format = bzfile(outfilename)
+  write.csv(csv, format)
+}
+
+# All of the zip file names
+compressed = c("hads2013n_ASCII.zip",
+               "hads2011(ASCII).zip",
+               "hads2009(ASCII)_v2.exe",
+               "hads2007(ASCII)_v2.exe",
+               "hads2005(ASCII)_v2.exe",
+               "hads2003(ASCII).exe",
+               "hads2001(ASCII).exe",
+               paste0("hads", seq(99,85,-2), "(ASCII).exe"))
+
+# All of the unzipped file names
+unzipped = c("thads2013n.txt",
+             str_c("thads", seq(2011,2005,-2), ".txt"),
+             str_c("hads", seq(2003,2001,-2), ".txt"),
+             str_c("hads", seq(99,85,-2), ".txt"))
+
+#Output names
+actualnames = str_c("hads", seq(2013,1985,-2), ".bz2")
+
+mapply(downloadfile, compressed, unzipped, actualnames)

### 2016-08-10 paper/Paper.Rmd
title changes

@@
 
-<br>
-<br>
-<br>
-
-#Introduction
-
-The dataset was taken from the American Housing Survey (AHS), which derives sets of data from 1985 and on.
-The data set includes categorizing each housing unit by factors such as affordability, income through fair market rent, burden, and the adjusted median income.
-Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
-
-  The goal is to determine whether housing affordability has increased or decreased over the last few decades.
-Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
-The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
-Another major event that has potential consequences on the current state of housing affordability, is the 2007 Housing Market crash.
-Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
+\vspace{72pt}
 
@@
   \newpage
+
+#Introduction
+
+The dataset was taken from the American Housing Survey (AHS), which derives sets of data from 1985 and on.
+The data set includes categorizing each housing unit by factors such as affordability, income through fair market rent, burden, and the adjusted median income.
+Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+
+  The goal is to determine whether housing affordability has increased or decreased over the last few decades.
+Our initial hypothesis assumes that housing affordability within the United States has decreased due to key, significant events that have occurred within the last few years.
+The United State's growing income inequality can be attributed to the current state of housing affordability, and is a clear indicator of the shifting income brackets between who is poor and who is rich.
+Another major event that has potential consequences on the current state of housing affordability, is the 2007 Housing Market crash.
+Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
@@
 
-#Graphical findings
+#Findings
 
@@
   coord_fixed(1.3) +
-  theme_tufte(base_size = 20, 
+  theme_tufte(base_size = 14, 
               base_family = "serif") +

### 2016-08-10 paper/Paper.Rmd
superscipts

@@
 Another major event that has potential consequences on the current state of housing affordability, is the 2007 Housing Market crash.
-Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
+According to Business Insider, the end of 2010 saw a staggering 23.1% of all U.S.
+homeowners with a mortgage owed more on their homes than what their homes were worth.
+^1^ Amongst all of the foreclosures, loan defaults, and stock market crash, we will closely examine the years between 2007 and 2011 to capture a better sense of the status of the housing market's affordability.
 
@@
 #Findings
+##Initial Findings
 
-```{r, echo = F, fig.cap="Median Burden Across Years and Household Sizes"}
+\vspace{24pt}
+
+```{r, echo = F,fig.cap="Median Burden Across Years and Household Sizes", fig.height=3}
+
 #Read in the cleaned data
-
 clean_years <- read_csv("../clean_data/clean_years.csv")
@@
 
+\vspace{12pt}
   Figure 1 shows the development of median burden (proportion of housing costs relative to household income) from 1985 to 2013.
@@
 In households of 3 or more people, the burden picks up again.
-This is likely due to a larger amount of children and / or retirees in larger households.
+This is likely due to a larger amount of children and/or retirees in larger households.
 Also, households with 3 or more members tend to hae only one earner.
@@
 
+\vspace{6pt}
   To map income levels in different areas we used the four census regions in the dataset and joined them with a scraped table from mapsoftheworld.com.
@@
 
-```{r, echo = FALSE, fig.width=3.5, fig.height=5, fig.cap='one plotthe other one'}
+```{r, echo = FALSE, fig.width=3.5, fig.height=4, fig.cap='one plotthe other one'}
 clean_years %>%
@@
   In order to better understand the regional differences in income, the incomes were adjusted by certain variables.
-
-  The first plot of figure 3 compares APLmed by year for the four regions in the United States (Northeast, South, Midwest, West).
+The first plot of Figure 3 compares the median income adjusted for number of persons by year (APLmed) for the four regions in the United States (Northeast, South, Midwest, West).
 APLmed is the adjusted median household income per person.
@@
 The reason for no adjustment to a four person household income is because the median household size in the United States is about four persons.
-
-  It is clear that the Northeast and the West are the most expensive regions.
+It is also clear that the Northeast and the West are the most expensive regions.
 This is makes sense given that the northeast and West are by far the most productive (by GDP) regions in the United States.
@@
 So, when the 2008 crisis occurred the repercussions were immediately felt in the Northeast (Wall Street) while the other regions took some time to correct for the downturn in the financial market.
-
-  The second plot of figure 3 depicts the development of median income adjusted for the number of bedrooms in a household's living quarters.
+The second plot of Figure 3 depicts the development of median income adjusted for the number of bedrooms in a household's living quarters.
 The adjusted median income was calculated using precaculated HUD indexes.
@@
 The lag in the downward trend after 2008 still occurs in the plots.
+
 With the income information in hand, the burden data can be explored in further detail.
 
-```{r, echo = FALSE, fig.cap = "Amount of households relative to poverty line"}
+##Further Exploration
+
+```{r, echo = FALSE, fig.cap = "Amount of Households Relative to the Poverty Line", fig.height = 3}
 clean_years %>% 
@@
 
-```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden"}
+```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden, fig.height = 1"}
 ###### Household Income
@@
 
-```{r, echo = FALSE, fig.cap = "Relationship between Monthly Housing Costs and Burden"}
+```{r, echo = FALSE, fig.cap = "Relationship between Monthly Housing Costs and Burden", fig.height = 3}
 
@@
 
+At a certain point, regardless of income, housing costs stay the same because housing prices have increased
+
+The no income line continues to have monthly housing costs due to people who lost their jobs, unemployment, trust fund recipients using their money, and people living off savings
+
   Figure 6 offers a more in-depth look at the relationship between housing costs and burden.
@@
 
-```{r, echo= FALSE, fig.cap="Relationship between Monthly Housing Costs and Burden"}
+```{r, echo= FALSE, fig.cap="Relationship between Monthly Housing Costs and Burden", fig.height = 3}
 
@@
 Regions could be further subsetted by town or district level to provide more granular data.
+
+\newpage
+
+#Sources
+
+1.
+Snyder, Michael.
+"27 Depressing Facts About The Housing Crash That Never Seems To End." Business Insider.
+2011.
+Accessed August 7, 2016.
+http://www.businessinsider.com/amazing-statistics-about-the-housing-crash-2011-3.
+
+2.

### 2016-08-10 paper/Paper.Rmd
final changes hopefully

@@
 The data set includes categorizing each housing unit by factors such as affordability, income through fair market rent, burden, and the adjusted median income.
-Certain factors, such as Poverty Income, is based on the Census Bureau's official thresholds.
+Certain factors, such as poverty income, is based on the Census Bureau's official thresholds.
 
@@
 
-```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden, fig.height = 1"}
+```{r, echo = FALSE, fig.cap="Relationship between Household Income and Burden, fig.height = 3"}
 ###### Household Income
@@
 
-  As pictured in figure 5, the relationship between household income and burden is seen as a positive correlation.
-From the graph, you can see that higher income household have less burden to deal with, which makes sense because those with more money are able to pay off monthly costs more easily.
-Those with 50% more burden are near the poverty line.
-Figure 5 also shows a spike in households that have burden.
-Households with higher incomes started to pay more burdens, which can be attributed to 2007 financial housing market crisis.
+  As pictured in Figure 5, the relationship between household income and burden is seen as a positive correlation.
+The relatively higher household income brackets are shown to have lower burdens than the lower household income brackets.
+Because burden is a proportion of housing cost over income, it is reasonable to conclude that higher income households will have a smaller burden due to their larger incomes.
+The higher income households are able to pay off their monthly costs more easily.
+Although the federal poverty line differs between different sized households, the average poverty threshold for families of 3 or 4 persons is ranged from $20k to $25K.
+^2^ From Figure 5, it is easily seen that those with 50% or More burden are near and/or underneath the poverty threshold.
+At the onset of the 21st century to 2010, there is a spike in 2007/2008 for the range of household income with the differing burden brackets.
+This spike can be attributed to the 2007 financial housing crisis, which ultimately affected the  burden levels of the housing sector.
 
@@
   summarise_(new_y = as.formula(
-    paste0("~ median(","zsmhc", ", na.rm=T)"))) %>%
+    paste0("~ median(",
+           "zsmhc", 
+           ", na.rm=T)"))) %>%
 
@@
 
-At a certain point, regardless of income, housing costs stay the same because housing prices have increased
-
-The no income line continues to have monthly housing costs due to people who lost their jobs, unemployment, trust fund recipients using their money, and people living off savings
-
   Figure 6 offers a more in-depth look at the relationship between housing costs and burden.
 It's interesting to see that all three burden lines are close to one another, which illustrates the principle that those with more money aren't buying more expensive houses.
-From figure 5 we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
-Each level of burden has similar housing costs, therefore reinforcing the idea that people are buying similar priced housing among the lower and middle classes.
-Again, a spike around 2010 is prevalent, which can be caused by the 2007 housing crisis.
+From Figure 5, we saw that those with higher incomes have less burden, and those with lower incomes have higher burden.
+At a certain point, regardless of income, housing costs stay the same.
+For the "No Income" bracket, there are still monthly costs present.
+To account for this anomly, we assum that people who lost their jobs could still be paying these costs with zero income.
+Trust fund recipients and people living off savings could also be paying off these costs.
+Again, a spike around 2010 is prevalent, which can be explained by the 2007 housing crisis.
 
@@
   cleanyearsmonthly %>%  
-  group_by_("year","region", "fmtburden") %>%  
+  group_by_("year",
+            "region", 
+            "fmtburden") %>%  
   summarise_(new_y = as.formula(
-    paste0("~ median(","zsmhc", ",na.rm=T)"))) %>% 
-  ggplot(aes_string(x = "year", y = "new_y", color = "fmtburden")) + 
+    paste0("~ median(",
+           "zsmhc", 
+           ",na.rm=T)"))) %>% 
+  ggplot(aes_string(x = "year", 
+                    y = "new_y", 
+                    color = "fmtburden")) + 
   geom_line(size = 1) + 
-  facet_wrap(~region, labeller = as_labeller(regionNames)) + 
-  labs(y = "Monthly Housing Costs", x = "Year", color = "Burden") +
-  scale_y_continuous(labels=function(x){paste0("$", x)}) +
+  facet_wrap(~region, 
+             labeller = as_labeller(regionNames)) + 
+  labs(y = "Monthly Housing Costs", 
+       x = "Year", 
+       color = "Burden") +
+  scale_y_continuous(labels = function(x){paste0("$", x)}) +
   scale_colour_brewer(palette = "Dark2") +
@@
 
-  From figure 7 you can see that the Northeast and West have higher rates of burden at higher income levels, therefore illustrating how housing affordability in these regions are less than that of the Midwest and South.
+  Figure 7 illustrates the monthly housing costs by census regions (Northeast, Midwest, South, and West).
+It is interesting to note that the Northeast and West have higher monthly housing costs according to the figure.
+According to Forbes, a majority of the most overpriced cities are located in these two regions.
+^3^ Figure 2 also supports this idea because median income is highest in these two regions.
+There is an interesting dip seen in the Northeast graph, which can be attributed to the Northeast being impacted earlier by the housing crisis.
 
@@
       # Use NSE
-      group_by_("year", "fmtstructuretype") %>%  
+      group_by_("year", 
+                "fmtstructuretype") %>%  
       # Take the median instead of mean to avoid extreme observations skewing the 
       # result.
-      summarise_(new_y= as.formula(paste0("~ median(","increlamipct", ",na.rm=T)"))) %>%  
+      summarise_(new_y = as.formula(
+        paste0("~ median(",
+               "increlamipct", 
+               ",na.rm=T)"))) %>%  
       # Use NSE
@@
 \begin{center}
-Figure 8: Income compared to area's median grouped by house type
+Figure 8: Income Compared to Area's Median Grouped by House Type
 \end{center}
@@
 
-```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median housing cost with assistance"}
+```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median Housing Cost with Assistance"}
 assisted = clean_years %>% filter(!is.na(assisted)) %>% 
@@
 
+ \newpage
+
 #Conclusion:
@@
 
-Ultimately, our findings point towards
+Ultimately, our findings point towards the idea that housing affordability has decreased since the mid 80s.
+Although many graphs illustrate a dip in burden, this does not negate the overarching idea that housing affordability has decreased.
+According to Investopedia, major consequences of the mortgage crisis include chronically higher unemployment and smaller chances of getting loans and credit.
+The housing bubble has had effects on a long-term scale, where in 2009, there was a 38% increase in Americans filing for bankruptcy since 2007.
+^4^ We conclude that the 2007 housing crisis is one of the most significant factors in causing the overall decrease in housing affordability.
 
@@
 
-2.
+2. "Poverty Guidelines." ASPE. 2015. Accessed August 3, 2016. https://aspe.hhs.gov/poverty-guidelines.
+
+3. "America's Most Overpriced Cities In 2015." Forbes. Forbes, 4 Mar. 2015. Web. 5 Aug. 2016. https://.forbes.com/sites/erincarlyle/.../americas-most-overpriced-cities-in-2015/.
+
+4. "5 Consequences Of The Mortgage Crisis." Investopedia. N.p., 21 Nov. 2010. Web. 9 Aug. 2016. http://investopedia.com/financial-edge/1110/5-consequences-of-the-mortgage-crisis.aspx

### 2016-08-10 clean_data/CleanData.R
Add quoted NAs to cleaning data

@@ -46,7 +46,7 @@ NoQuotations <- function(x) {
   str_replace_all(x, "^\\'|\\'$", "")
 }
 # Total list of invalid values, done due to insufficient documentation.
-NA_LIST = c("","NA","-5","-6","-7","-9",".")
+NA_LIST = c("", "NA", "-5", "-6", "-7", "-9", ".", "''", "'NA'", "'-5'", "'-6'", "'-7'", "'-9'", "'.'")
 # Loading the most recent datasets first to give reference for other
 # two sets.
 base_files <- paste("../raw_data/hads",          

### 2016-08-10 paper/Paper.Rmd
final changes

@@
 \begin{center}
-Figure 7: Relationship between Monthly Housing Costs and Burden in different regions
+Figure 7: Relationship between Monthly Housing Costs and Burden in Different Regions
 \end{center}
+
+\vspace{48pt}
 
@@
 
-```{r, fig.width=12, fig.height=6, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median Housing Cost with Assistance"}
+```{r, fig.width=15, fig.height=10, echo = F, message=F,error=F, warning=FALSE, fig.cap="Burden based on Median Housing Cost with Assistance"}
 assisted = clean_years %>% filter(!is.na(assisted)) %>% 
       # Use NSE
-      group_by_("assisted", "fmtcostmedrelamicat") %>%  
+      group_by_("assisted", 
+                "fmtcostmedrelamicat") %>%  
       # Take the median instead of mean to avoid extreme observations skewing the 
       # result.
-      summarise_(new_y= as.formula(paste0("~ median(","burden", ",na.rm=T)"))) %>% 
+      summarise_(new_y= as.formula(
+        paste0("~ median(",
+               "burden", 
+               ",na.rm=T)"))) %>% 
       ungroup %>% 
       mutate(assisted = as.factor(assisted))
-ggplot(assisted, aes(y = new_y,   
+ggplot(assisted, 
+       aes(y = new_y, 
            x = fmtcostmedrelamicat,
@@
        fill="Assistance") + 
-  scale_y_continuous(limits=c(0,1), labels=function(x){paste0(x*100, "%")}) +
-  scale_x_discrete(labels = function(x){str_replace_all(str_replace_all(x, "^[0-9] (.*)","\\1"), "LTE (.*)","<= \\1")}) +
+  scale_y_continuous(limits=c(0,1), 
+                     labels=function(x){paste0(x*100, "%")}) +
+  scale_x_discrete(labels = function(x){str_replace_all(
+    str_replace_all(x,
+                    "^[0-9] (.*)","\\1"), 
+    "LTE (.*)","<= \\1")}) +
   scale_fill_brewer("Assistance", 
@@
   coord_flip() +
-  theme_minimal(base_family = "serif")
+  theme_minimal(base_family = "serif", base_size = 15)
 ```
